### 2019-02-19 paper/bin-love-after-gw170817.tex

\documentclass[prd,twocolumn,nofootinbib,superscriptaddress,amsmath,amssymb]{revtex4-1}
\include{preamble}

\begin{document}

\title{Equation-of-state insensitive relations after GW170817}

\author{Zachary Carson}
\affiliation{%
 Department of Physics, University of Virginia, Charlottesville, Virginia 22904, USA
}%

\author{Katerina Chatziioannou}
%\affiliation{%
% Canadian Institute for Theoretical Astrophysics, 60 St. George Street, Toronto, Ontario, M5S 3H8, Canada
%}%
\affiliation{%
 Center for Computational Astrophysics, Flatiron Institute, 162 5th Ave, New York, NY 10010
}%

\author{Carl-Johan Haster}
\affiliation{%
 Canadian Institute for Theoretical Astrophysics, 60 St. George Street, Toronto, Ontario, M5S 3H8, Canada
}%
\affiliation{%
 Center for Computational Astrophysics, Flatiron Institute, 162 5th Ave, New York, NY 10010
}%

\author{Nicol\'as Yunes}
\affiliation{%
eXtreme Gravity Institute, Department of Physics, Montana State University, Bozeman, MT 59717, USA.
}%
\affiliation{%
 Department of Physics and MIT Kavli Institute,
}%

\author{Kent Yagi}
\affiliation{%
 Department of Physics, University of Virginia, Charlottesville, Virginia 22904, USA
}%

\date{\today}

%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%
%%%% ABSTRACT %%%%
%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%

\begin{abstract}
% Intro
The thermodynamic relation between pressure and density (i.e.~the equation of state) of cold supranuclear matter is critical to neutron stars, yet it remains one of the largest uncertainties in nuclear physics. 
% How to probe
The extraction of tidal deformabilities from the gravitational waves emitted in the coalescence of neutron star binaries, such as GW170817, is a promising tool to probe this thermodynamic relation.
% What was done before - Binary Love
Equation-of-state insensitive relations between symmetric and antisymmetric combinations of individual tidal deformabilities, the so-called  ``binary Love relations", have proven important to infer the radius of neutron stars, and thus constrain the equation of state, from such gravitational waves. 
% What was done before - I-Love-Q
A similar set of relations between the moment of inertia, the tidal deformability, the quadrupole moment, and the compactness of neutron stars, the so-called ``I-Love-Q" and ``C-Love" relations, allow for future tests of General Relativity in the extreme gravity regime. 
% What we will do
But even the most insensitive of such relations still presents some degree of equation-of-state variability that could introduce systematic uncertainties in parameter extraction and in model selection. 
%
We here reduce this variability by more than $50\%$ by imposing a prior on the allowed set of equations of state, derived from the posteriors generated from the analysis of GW170817.  
%
The resulting increase in insensitivity reduces systematic uncertainties in the extraction of the tidal deformability from future gravitational wave observations, although statistical uncertainties currently dominate the error budget, and will continue to do so until the era of Voyager-class detectors.  
\end{abstract}

\maketitle

%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%
%%%% INTRODUCTION %%%%
%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%

\section{Introduction}
\label{sec:intro}

%Intro
The thermodynamic relation between pressure and density in cold, supranuclear matter, the so-called equation of state (EoS), remains an unsolved problem in nuclear physics. The EoS is critical to our understanding of neutron stars (NSs) because it determines many NS observables, such as their mass, radius, moment of inertia (I), quadrupole moment (Q) and tidal deformability (or Love number). Unfortunately, terrestrial experiments can only probe the EoS to around nuclear saturation density ($\rho_{\text{sat}} \approx 2.5 \times 10^{14} \text{ g/cm}^3$)~\cite{Li:HeavyIon,Tsang:SymmetryEnergy,Centelles:NeutronSkin,Li:CrossSections,Chen:SymEnergy}. Although some temperature-dependent heavy-ion collision experiments are beginning to probe higher densities~\cite{Danielewicz:2002pu}, astrophysical observations of NSs remain ideal for constraining the EoS of cold and ultra dense, nuclear matter.

%Constraining the EoS
Independent measurements of NS observables can be used to constrain the nuclear EoS. For example, electromagnetic observations of the mass and radius of certain NSs have been used to place confidence limits in the mass-radius plane, and thus constrain the EoS~\cite{guver,ozel-baym-guver,steiner-lattimer-brown,Lattimer2014,Ozel:2016oaf}. These observations, however, may potentially suffer from large systematic errors due to uncertainties in the astrophysical modeling of X-ray bursts. The gravitational waves (GWs) emitted in the coalesce of NS binaries may be a cleaner probe of nuclear physics. During the early inspiral, the orbital separation is large enough that the tidal fields are negligible; but as the orbital separation decreases due to GW emission, tidal forces grow and the NSs respond by developing deformations controlled by their nuclear EoS. These deformations affect the orbital trajectory of the binary, and thus the GWs emitted, encoding in the latter the NS EoS~\cite{hinderer-love,Flanagan2008}.

%reparameterization of the template
The GWs emitted by binary NSs in the late inspiral must then depend on the \emph{tidal deformabilities} $\Lambda_1$ and $\Lambda_2$, which control the linear response of the star's quadrupole deformation to the (electric-type, quadrupole) tidal field of the companion (to leading order in a post-Newtonian expansion~\cite{Blanchet:2013haa})~\cite{Flanagan2008,Vines:2011ud}. These parameters, however, enter the GW Fourier model multiplied by the same function of the GW frequency (at leading post-Newtonian order), making them degenerate, and thus, very difficult to estimate independently with current GW data~\cite{Wade:tidalCorrections}. Instead, one can extract certain combinations of the tidal deformabilities, like the so-called \emph{chirp deformability} $\tilde{\Lambda}$~\cite{Favata:2013rwa,Wade:tidalCorrections}, or one can extract the (mass-independent) coefficients $(\lambda_{0},\lambda_{1},\ldots)$ of a Taylor expansion of the tidal deformabilities about some fiducial mass~\cite{delPozzo:TaylorTidal,Yagi:binLove}. Current detectors are not sensitive enough to accurately measure any of these coefficients, but future detectors will and the information from multiple events could then be combined, since the Taylor expansion coefficient should be common to all events.

%Universal relations
Lacking enough sensitivity in current GW observations, one is forced to only estimate the chirp deformability, but this prevents the independent extraction of $\Lambda_{1}$ and $\Lambda_{2}$. Yagi and Yunes~\cite{Yagi:binLove} solved this problem by finding ``approximately universal'' or ``EoS-insensitive'' relations between the symmetric and anti-symmetric combinations of the tidal deformabilities $\Lambda_{s,a}=\frac{1}{2}(\Lambda_1 \pm \Lambda_2)$, the so-called ``binary Love relations." These relations can be used to analytically express $\Lambda_{s}$ in terms of $\Lambda_{a}$ (or vice-versa), making the chirp deformability a function of only $\Lambda_{a}$. A measurement of the chirp deformability then implies a measurement of $\Lambda_{a}$, and through the use of the binary Love relations, also a measurement of $\Lambda_{s}$, which then allows for the inference of the individual tidal deformabilities $\Lambda_{1}$ and $\Lambda_{2}$~\cite{Yagi:binLove}. With those at hand, one can further use EoS-insensitive relations between the tidal deformabilities and the compactness, the so-called ``C-Love relations''~\cite{Yagi:ILQ}, to infer the radii of the NSs, and thus, to place two constraints in the mass-radius plane, one for each star in the binary~\cite{Yagi:binLove}. This idea was recently implemented in GW170817, allowing for the first EoS-independent constraints on the mass-radius curve using GW data~\cite{LIGO:posterior}.

EoS-insensitive relations can in fact be used for more than just measuring the nuclear EoS. For years, the theoretical physics community considered the possibility of using measurements of NS properties, like the mass, the radius and the moment of inertia, to constrain deviations from General Relativity in the strong-field regime. Certain modified theories of gravity, like scalar tensor theories with spontaneous scalarization~\cite{Damour:1996ke}, Einstein-Aether and Horava gravity~\cite{Eling:2007xh,Yagi:2013ava,Yagi:2013qpa}, dynamical Chern-Simons gravity~\cite{Yunes:2009ch}, beyond Horndesky theories~\cite{Sakstein:2016oel}, modify such NS observables, but unfortunately, these modifications are typically degenerate with the nuclear EoS. Yagi and Yunes solved this problem by finding EoS-insensitive relations between the moment of inertia, the tidal deformability (or Love number) and the quadrupole moment, the so-called ``I-Love-Q'' relations~\cite{Yagi:ILQ}. Given a measurement of the Love number for a given NS, for example through GW observations, the I-Love-Q relations can be used to infer the moment of inertia or the quadrupole moment. A second independent measurement of either of these two quantities, for example through binary pulsar observations~\cite{Lattimer:2004nj} or observations with the Neutron star Interior Composition ExploreR (NICER)~\cite{Ozel:2015ykl}, then allows an EoS-insensitive test of General Relativity in the strong field regime~\cite{Yagi:ILQ}.

The implementation of the EoS-insensitive relations in data analysis has to somehow contend with the fact that these relations are in fact not exactly universal, but rather present different (albeit small) levels of EoS variability. In the case of the binary Love and C-Love relations to infer the radii of NSs with GW170817, the problem is solved by marginalizing over the EoS variability~\cite{Katerina:residuals}. Presumably, this same procedure can be applied in the future when carrying out tests of General Relativity with the I-Love-Q relations, using a combination of binary pulsar, NICER and GW data. But this marginalization procedure may not always be as important; as constraints in the mass-radius plane become more stringent with future GW observations, the allowed space of EoSs will shrink, which in turn must naturally decrease the degree of EoS variability in all EoS-insensitive relations. This is the main focus of this paper.

%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%% EXECUTIVE SUMMARY %%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Executive Summary}

%How we improved this
We begin by investigating the increase in EoS insensitivity due to the constraints placed by GW170817 on the allowed space of EoSs. We take the 90\% credible region constructed from the posterior probability distributions on the pressure-density plane~\cite{LIGO:posterior} obtained from GW170817~\cite{TheLIGOScientific:2017qsa} as a prior on the set of EoSs that is compatible with this observation. We then generate two large samples of spectral EoSs~\cite{Lindblom:2018rfr}, one in which this prior is enforced (the ``constrained'' sample) and another in which the prior is not enforced (the ``unconstrained'' sample). We repeat the analysis done by Yagi and Yunes~\cite{Yagi:binLove,Yagi:ILQ} on both sets of EoSs and find that the EoS-insensitive relations present less EoS variability with the constrained set. In particular, the EoS-insensitivity increases by a factor of $\sim 60$\% in the binary Love relations (for stars with mass ratio larger than $0.75$), by a factor of $\sim 70\%$ in the C-Love relations, and by factors of $\sim 50$\% in the I-Love-Q relations (see Table~\ref{tab:maxVar} and Fig.~\ref{fig:ILQ} for more details).

With this study at hand, we then carry out additional related studies on EoS-insensitive relations that go beyond the work in~\cite{Yagi:binLove,Yagi:ILQ}. First, we investigate the relation between the NS radius and its tidal deformability, the R-Love relations, for both sets of EoSs, because these are critical to place constraints on the radius from a measurement of the Love number.  We use the C-Love relations to construct the R-Love relations and find that the maximum EoS variability drops from $\sim 880 {\rm{km}}$ in the unconstrained case to $\sim 360 {\rm{km}}$ in the constrained case (see Sec.~\ref{sec:clove} for more details). Second, we study the EoS-universality of hybrid stars, which experience strong first-order phase transitions from hadronic to quark matter in the core~\cite{Paschalidis2018}. We find that the I-Love-Q and C-Love relations remain EoS-insensitive for these hybrid stars, although the EoS variability increases slightly (see Sec.~\ref{sec:ilq-hyb} and~\ref{sec:clove-hyb} for more details). However, we also find that the binary Love relations are not EoS-insensitive for a mixed binary with a (massive) hybrid star and a (low-mass) hadronic star, due to the large separation in chirp deformability between the constituent stars (see Sec.~\ref{sec:binary} for more details).

Last for not least, we study the importance of using the improved EoS-insensitive relations in future GW observations. The use of EoS-insensitive relations introduces systematic uncertainties in parameter estimation because of the intrinsic non-zero EoS-variability in these relations, which one must marginalize over. These uncertainties are currently irrelevant because statistical uncertainties in parameter estimation are much larger with current detectors. But as the detector sensitivity is improved, the signal-to-noise ratio and the number of events that will be detected will increase, therefore decreasing the statistical uncertainties below the systematic one due to EoS variability. We carry out a simple Fisher analyses to estimate when the statistical uncertainties become comparable to the systematic uncertainties due to EoS variability and find that this occurs for Voyager-class detectors.

Figure~\ref{fig:stackedFisher} shows this result in more detail. We here present the (Fisher-estimated) statistical uncertainty in the measurement of $\lambda_{0}$ with various detectors (LIGO O2~\cite{aLIGO}, Advanced LIGO (aLIGO)~\cite{aLIGO}, LIGO A\texttt{+} (A\texttt{+})\cite{Ap_Voyager_CE}, Voyager~\cite{Ap_Voyager_CE}, the Einstein Telescope (ET)~\cite{ET}, and the Cosmic Explorer (CE)~\cite{Ap_Voyager_CE}) for an event identical to GW170817. The x-axis shows the signal-to-noise ratio for a GW170817 event detected with each of these detectors. We also present the combined statistical uncertainty $\sigma^A_N$ after $N$ detections with each of these instruments, with the top and bottom of the region representing the most optimistic and pessimistic expectation for the number of detections expected from the binary NS merger rate (see also Table~\ref{tab:variances}). These statistical uncertainties should be compared to the systematic uncertainty in $\lambda_{0}$ due to EoS-variability improved with the constrained set. Observe that the statistical and the systematic uncertainties cross for Voyager-class detectors.

\begin{figure}
\begin{center} 
\includegraphics[width=\columnwidth]{stackedFisher.eps}
\end{center}
\caption{(Color online) Fisher-estimated statistical uncertainties on the extraction of $\lambda_{0}$ with interferometer (O2, aLIGO, A\texttt{+}, Voyager, CE, ET-D) as a function of the signal-to-noise-ratio expected in each of these instruments, given a single GW170817 detection (circles). The statistical uncertainties with ET are lower than with CE in spite of a lower signal-to-noise ratio because the former is more sensitive above 300 Hz, where tidal effects matter the most (see Sec.~\ref{sec:futureObservations} for further discussion). We also plot the combined statistical uncertainty given $N$ observations consistent with the NS binary merger rate for a 1 year observation (regions), with the top and bottom edges of the regions corresponding to optimistic and pessimistic merger rates. These statistical uncertainties should be compared to the systematic uncertainty on the extraction of $\lambda_{0}$ due to EoS-variability (horizontal dashed line). Observe that the statistical and systematic uncertainties cross for Voyager-class detectors. We confirm this conclusion by repeating the statistical analysis with two different waveform models (PhenomD~\cite{PhenomDI,PhenomDII} plus NRTidal corrections~\cite{Samajdar:NRTidal} and PhenomD~\cite{PhenomDI,PhenomDII} plus 6PN tidal corrections~\cite{Wade:tidalCorrections}).}
\label{fig:stackedFisher}
\end{figure}

%outline
The remainder of this paper presents the details of the results summarized above and it is organized as follows. 
We begin with complementary background and theory material in Sec.~\ref{sec:theory}.
We continue in Sec.~\ref{sec:universal} by finding new and improved binary Love, I-Love-Q, and C-Love relations, and considering how well hybrid star EoSs agree with the EoS-insensitive relations.
We next examine these improved relations and question whether or not they are useful for future interferometers in Sec.~\ref{sec:observations}.
We conclude in Sec.~\ref{sec:conclusion} by discussing our results and mentioning avenues of future work.
Throughout this paper, we have adopted geometric units of $G=1=c$, unless otherwise stated.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%% BACKGROUND AND THEORY %%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Background and theory}\label{sec:theory}

In this section we review how the EoS can be represented analytically
through a spectral decomposition, and how the observation of GW170817
constrains the space of possible EoSs. We then proceed to discuss 
how one computes the tidal deformabilities and the EoS-insensitive relations.

\subsection{Spectral representations of NS equations of state}
\label{sec:eos}

The structure of a NS and its tidal interactions in a binary system rely heavily on the underlying state function (or equation of state - EoS) describing the relationship between the pressure ($p$) and energy density ($\epsilon$) of nuclear matter.
Given that all currently proposed EoSs utilize certain approximations~\cite{Oertel:Review,Baym:Review}, one method to study a wide range of physically realizable EoSs is to parameterize them such that any realistic EoS can be represented with a small number of parameters.
Spectral representations~\cite{Lindblom:2010bb,Lindblom:2012zi,Lindblom:2013kra,Lindblom:2018rfr,Abbott:2018exr} parameterize EoSs by performing spectral expansions on the adiabatic index $\Gamma(p)$\footnote{Another way of parameterizing EoSs is through a piecewise polytropic formulation~\cite{Read2009,Lackey:2014fwa,Carney:2018sdv}.}:
\begin{equation}
\Gamma(x) = \exp{\sum_k^{N}\gamma_k x^k},
\end{equation}
where $x \equiv \log{(p/p_0)}$ for a minimum pressure $p_0$.
The equation of state is then determined by an integration of the differential equation:
\begin{equation}
\frac{d \epsilon(p)}{dp}=\frac{\epsilon(p)+p}{p \Gamma(p)}.
\end{equation}
Using this formalism, any valid EoS can be approximated through the choice of $N$ spectral coefficients $\gamma_k$, and we here choose $N=4$, tabulated for several common EoSs in Table 1 of~\cite{Lindblom:2018rfr}.

We here wish to consider EoS-insensitive relations using two sets of EoSs: a ``constrained set'' consistent with the observation of GW170817 and an ``unconstrained set'' that does not impose this prior. In \emph{both} cases, we model the EoS with a $C^{0}$-piecewise function that equals the low-density crust EoS of SLy~\cite{Douchin:2001sv} below half nuclear saturation density $\rho_{\text{stitch}}=1.3 \times 10^{14} \text{ g/cm}^3$~\cite{Read2009}, and equals the spectral decomposition described above outside the crust. For the latter, we restrict the spectral coefficients to the ranges $\gamma_0 \in \lbrack 0.2,2 \rbrack$, $\gamma_1 \in \lbrack -1.6,1.7 \rbrack$, $\gamma_2 \in \lbrack -0.6,0.6 \rbrack$, $\gamma_3 \in \lbrack -0.02,0.02 \rbrack$, and the adiabatic index is further restricted to $\Gamma \in \lbrack 0.6,4.5 \rbrack$~\cite{Lindblom:parameters}. Moreover, we impose the following two restrictions:  (i) causality within 10\%, i.e. that the speed of sound of the fluid be less than the speed of light to 10\%, and (ii) a high maximum mass, i.e.~that the resulting EoS supports NSs with masses at least as high as $1.97 \text{ M}_{\odot}$, consistent with astrophysical observations~\cite{Zhao:massiveNS}.

The unconstrained set is then defined by drawing random samples in the spectral coefficients (within their allowed prior ranges), and then eliminating any EoS that either leads to a polytropic index $\Gamma$ outside the allowed range, breaks the causality restriction, or breaks the maximum mass restriction. The constrained set is defined in a similar way, with the addition of the restriction that the resulting EoS be inside the 90\% marginalized posterior distribution on the pressure-mass density plane derived from the GW170817 observation~\cite{LIGO:posterior,Carney:2018sdv}  (see Fig.~2 in~\cite{LIGO:posterior}). In both cases, each set consists of 100 members, with a subset of these shown in the left panel of Fig.~\ref{fig:eos}.

%In the current analysis, we consider EoSs which have been constrained by the recent binary neutron star merger event, GW170817.
%Important recent work by the LIGO Collaboration~\cite{LIGO:posterior,Carney:2018sdv} sampled the EoS parameter space in order to derive a marginalized posterior on the pressure as a function of mass density, as seen in Fig. 2 of~\cite{LIGO:posterior}.
%The spectral coefficients were sampled within the ranges: $\gamma_0 \in \lbrack 0.2,2 \rbrack$, $\gamma_1 \in \lbrack -1.6,1.7 \rbrack$, $\gamma_2 \in \lbrack -0.6,0.6 \rbrack$, $\gamma_3 \in \lbrack -0.02,0.02 \rbrack$, and further restricted the adiabatic index to be $\Gamma \in \lbrack 0.6,4.5 \rbrack$, ensuring the parameterization exposed a wide range of viable EoSs~\cite{Lindblom:parameters}.
%Additional constraints imposed upon the generated EoSs were as follows~\cite{LIGO:posterior}: (i) causality within 10\%, and (ii) EoS priors must support NS masses up to $1.97 \text{ M}_{\odot}$, consistent with astrophysical observations~\cite{Zhao:massiveNS}.
%We utilize a random set of 100 of the above posterior samples defined by GW170817, furthermore referenced as the ``constrained EoSs", as shown in Fig.~\ref{fig:eos}.
%For comparison, we also include a second sample of 100 ``unconstrained" EoSs, randomly sampled in the pressure-density plane to be unrestricted by GW170817s posterior.
%Following~\cite{Read2009}, the parameterized high-density core EoSs generated above are matched to the low-density crust EoS of Sly~\cite{Douchin:2001sv} at about half of the nuclear saturation density, $\rho_{\text{stitch}}=1.3 \times 10^{14} \text{ g/cm}^3$.

In addition, we investigate 10 transitional quark-hadron matter stars, which undergo strong first-order phase transitions at a pressure $P_{\text{tr}}$, leading to the hadronic branch departing into a quark-matter branch at a given transitional mass~\cite{Paschalidis2018,Alford:2017qgh,1971SvA....15..347S,Zdunik:2012dj,Alford:2013aca}. In particular, we focus on the ACS and ACB models described in~\cite{Paschalidis2018}, and also shown in the right panel of Fig.~\ref{fig:eos}. These result in two distinct types of NSs, based on their mass: (i) massive ($M \geq \text{ M}_{\text{tr}}$) hybrid stars which have quark-matter inner cores and nuclear matter elsewhere (henceforth, we denote such stars as hybrid stars (HSs)), and (ii) low-mass ($M \leq \text{ M}_{\text{tr}}$) hadronic stars with no internal transition to quark matter (henceforth, we denote these stars as simply NSs).

\begin{figure*}
\begin{center} 
\includegraphics[width=\columnwidth]{EoSs.eps}
\includegraphics[width=\columnwidth]{hybridEoSs.eps}
\end{center}
\caption{(Color Online) (left) Small representative samples of the unconstrained (dotted) and constrained (solid) sets, together with the 90\% marginalized posterior distribution from the observation of GW170817 (cyan shaded region)~\cite{LIGO:posterior}. Observe that the there is significantly less variability in the constrained set of EoSs due to the requirement that they be consistent with the GW170817 observation.
(right) ACS and ACB model hybrid stars investigated in our analysis, each transitioning from a hadronic branch (corresponging to a pure hadronic-matter NS) into a quark-matter branch (quark-matter inner core surrounded by hadronic matter) at various transition pressures $P_{\text{tr}}$.
}
\label{fig:eos}
\end{figure*}

\subsection{NS Tidal deformability}\label{tidal}

The GWs emitted in the coalescence of binary NSs, such as GW170817, provide valuable insight into their internal structure. As the stars spiral into each other, they tidally deform due to their companion's tidal field. The tidal field can be decomposed into multipole moments, with the leading-order field being even parity and quadrupolar, and thus leading to an even parity and quadrupolar deformation in response. The constant of proportionality that determines how quadrupolarly deformed a star becomes in the presence of a quadrupolar tidal field is called the \emph{tidal deformability} and denoted $\lambda$. Therefore, this parameter has the largest impact on the GW phase, and thus, it is encoded in GW observations of binary NS coalescence.

Consider a NS of mass $M$ in the presence of the (spatial) tidal tensor $\varepsilon_{ij}$ of its companion. In response to this tidal field, the NS will deform away from sphericity and acquire a (spatial) quadrupole moment tensor $Q_{ij}$. The tidal deformability $\lambda$ quantifies this linear response via~\cite{Flanagan2008,hinderer-love,Yagi2013}:
\begin{equation}
Q_{ij}=-\lambda \; \varepsilon_{ij}.
\end{equation}
One can extract $Q_{ij}$ and $\varepsilon_{ij}$ -- and thus $\lambda$, or its dimensionless form $\Lambda \equiv \lambda/M^5$ -- via a double asymptotic expansion of the gravitational potential in a buffer zone, defined by ${\cal{L}} \gg r \gg R$:
\begin{align}
\Phi=\frac{1+g_{tt}}{2}&=-\frac{M}{r} - \frac{3}{2}\frac{Q_{ij}}{r^3} \Bigg(\frac{x^i}{r} \frac{x^j}{r}-\frac{1}{3}\delta_{ij} \Bigg) + \mathcal{O} \Bigg( \frac{{\cal{L}}^4}{r^4} \Bigg)
\nonumber \\ &+ \frac{1}{2} \varepsilon_{ij} x^i x^j + \mathcal{O} \Bigg( \frac{r^3}{R^3} \Bigg),
\end{align}
where $R$ is the radius of the star and ${\cal{L}}$ is the length scale associated with the radius of curvature induced by the companion.

The calculation of the tidal deformability therefore requires the calculation of the metric component $g_{tt}$. One typically starts by first constructing a spherically symmetric, non-spinning background solution, whose stellar mass and radius are determined from $p(R)=0$. One then perturbatively introduces a tidal deformation and solves the perturbed Einstein equations in the NS interior, and matches this solution to an exterior Schwarzschild solution at the surface, modulo a constant. The requirement that the metric be at least $C^{1}$ fixes all constants of integration, yielding the tidal deformability, as discussed in more detail in~\cite{hinderer-love}. One can similarly obtain the moment of inertia from the asymptotic behavior of the $g_{t \phi}$ metric component.

Consider now a binary NS system, as that which produced the GW170817 event, where each star individually experiences the tidal field of the other star.
Thus, each star possesses tidal deformabilities $\Lambda_1$ and $\Lambda_2$ that enter the GW phase and encode information about each star's EoS.
Due to strong correlations between these parameters, individual extraction is very difficult with current interferometer sensitivity limitations.
However, the waveform templates can be strategically reparameterized to instead include independent linear combinations of $\Lambda_1$ and $\Lambda_2$ that mitigate correlations. One common reparameterization is through the introduction of the chirp deformability $\tilde{\Lambda}=\tilde{\Lambda}(\Lambda_1,\Lambda_2)$ and the parameter $\delta \tilde{\Lambda}=\delta \tilde{\Lambda}(\Lambda_1,\Lambda_2)$~\cite{Wade:tidalCorrections}. Since these parameters enter the GW phase at 5 and 6PN orders respectively\footnote{A term of NPN order is proportional to $v^{2N}$ relative to the leading-order term in the expression.}, they partially break the degeneracies between $\Lambda_{1}$ and $\Lambda_{2}$.

\subsection{EoS-insensitive relations}\label{sec:eosInsensitive}
Current GW interferometry is not yet sensitive enough to accurately extract both tidal parameters $\tilde{\Lambda}$ and $\delta\tilde{\Lambda}$.
In a search to remedy this, Yagi and Yunes~\cite{Yagi:binLove} found that symmetric and antisymmetric combinations of the tidal deformabilities:
\begin{equation}
\Lambda_s \equiv \frac{\Lambda_1 + \Lambda_2}{2}, \hspace{6mm} \Lambda_a \equiv \frac{\Lambda_1 - \Lambda_2}{2},
\end{equation}
display EoS-insensitive properties to a high degree, showing EoS variations of at most 20\% for binaries with masses less than $1.7 \text{ M}_{\odot}$ and using a representative sample of 11 EoSs. 
These ``binary Love relations" allow one to analytically break degeneracies between the tidal parameters: one can substitute $\Lambda_{a}=\Lambda_{a}(\Lambda_{s})$ in the GW model, thus completing eliminating $\Lambda_{a}$ from the parameter list (or vice-versa). This is important for two reasons: (i) the new model allows for the more accurate extraction of $\Lambda_{s}$ (or $\Lambda_{a}$ if $\Lambda_{s}$ is eliminated), and (ii) the relations allow for the inference of $\Lambda_{a}$ given a measurement of $\Lambda_{s}$ (or vice-versa), and from this for the inference of the individual tidal deformabilities $\Lambda_{1}$ and $\Lambda_{2}$. A simple Fisher analysis has shown that the binary Love relations improve parameter estimation of $\tilde{\Lambda}$ by up to an order of magnitude~\cite{Yagi:binLove}.

Similar EoS-insensitive relations have been found between individual NS observables: the moment of inertia (I), the tidal deformability (Love), the quadrupole moment (Q), and the compactness (C), known as the ``I-Love-Q" and ``C-Love" relations~\cite{Yagi:ILQ, Yagi:binLove}.
These relations are EoS-insensitive to better than 1\% and 6\% respectively, and they have been important applications in both GW astrophysics and experimental relativity. For example, these relations and the measurement of the tidal deformabilities allow for the inference of several other stellar properties, such as the moment of inertia, the  compactness, the spin ($\chi$), and the radius of NSs~\cite{Kumar:2019xgp}.
Analyses such as that of~\cite{Kumar:2019xgp} could benefit from the improvement of such EoS-insensitive relations, which we derive in this paper.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%% UNIVERSAL RELATIONS %%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{EoS-insensitive relations}
\label{sec:universal}

In this section, we repeat and improve the analyses of~\cite{Yagi:binLove,Yagi:ILQ} through the use of the two sets of EoSs (the constrained and the unconstrained set) and through new fitting functions that properly limit Newtonian results.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%% I-LOVE-Q RELATIONS %%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{I-Love-Q relations}
\label{sec:ilq}

Here we present our results on the I-Love-Q universality, comparing when possible to Fig.~1 of~\cite{Yagi:ILQ}.
In particular, we consider two distinct classes of NSs: nuclear matter EoSs and hybrid quark-hadron star EoSs as described in Sec.~\ref{sec:theory}.
We begin in Sec.~\ref{sec:ilq-nuc} by fitting the new I-Love-Q relations using the constrained set of EoSs.
This is followed in Sec.~\ref{sec:ilq-hyb} by an analysis and discussion of how well the hybrid star EoSs agree with the improved binary Love relations.

\subsubsection{Nuclear matter stars}
\label{sec:ilq-nuc}

Following~\cite{Yagi:ILQ}, we first fit the data for each EoS-insensitive relation to the following function:
\begin{equation}\label{eq:ILQfit}
\ln{y_i}=a_i+b_i \ln{x_i} + c_i (\ln{x_i})^2 + d_i (\ln{x_i})^3 + e_i (\ln{x_i})^4,
\end{equation}
where $y$ and $x$ correspond to NS observables $\bar{I}$, $\bar{Q}$, and $\Lambda$, and the updated coefficients are given in Table~\ref{tab:ILQfit}. This fitting function, however, does not limit properly to the Newtonian results~\cite{Yagi:ILQ}:
\begin{equation}\label{eq:Newtonian}
\bar{I}^{\text{N}} = K_{\bar{I}\Lambda}\Lambda^{2/5}, \hspace{3mm} \bar{Q}^{\text{N}} = K_{\bar{Q}\Lambda}\Lambda^{1/5}, \hspace{3mm} \bar{I}^{\text{N}} = K_{\bar{I}\bar{Q}}\bar{Q}^{2},
\end{equation}
where $\Lambda^{-1/5} \sim C$ when $C \ll 1$. One can thus improve the fitting function to 
\begin{equation}\label{eq:ILQfitNew}
y=K_{yx} x^{\alpha} \frac{1+\sum_{i=1}^3 a_i x^{-i/5}}{1+\sum_{i=1}^3 b_i x^{-i/5}},
\end{equation}
where $\alpha$ is either $2/5$, $1/5$, or $2$ for the $\bar{I}-\Lambda$, $\bar{Q}-\Lambda$, and $\bar{I}-\bar{Q}$ relations respectively, and where the new fitting coefficients are presented in Table~\ref{tab:ILQfitNew}. While the two fits result in similar $R^2$ values\footnote{$R^2$ is the coefficient of determination, defined as $\sum_i(f_i-\bar{y})^2/\sum_i(y_i-\bar{y})^2$, where $\bar{y}$ is the mean data value, and $f_i$, $y_i$ are the modeled and actual data values} of $\sim 0.999995$, the fit in Eq.~\eqref{eq:ILQfitNew} has the advantage that it properly limits to the Newtonian result as $\Lambda \gg 1$~\cite{Yagi:binLove}.

{\ny{Too many tables. Please combine Tables 1 and 2 into a single table and change the notation of the fits appropriately.}}
\begin{table*}
\centering
\caption{
Updated fit parameters for the I-Love-Q relations, fitted to the constrained EoS data by the curve found in Eq.~\ref{eq:ILQfit}.
}\label{tab:ILQfit}
\begin{tabular}{ c  c | c c c c c } 
 \hline
 \hline
 $y_i$ & $x_i$ & $a_i$ & $b_i$ & $c_i$ & $d_i$ & $e_i$ \\
 \hline
 $\bar{I}$ & $\Lambda$ & $1.493$ & $0.06410$ & $0.02085$ & $-5.018 \times 10^{-4}$ & $3.16 \times 10^{-7}$ \\
 $\bar{Q}$ & $\Lambda$ & $0.2093$ & $0.07404$ & $-0.05382$ & $-5.018 \times 10^{-3}$ & $1.576 \times 10^{-4}$ \\ 
  $\bar{I}$ & $\bar{Q}$ & $1.383$ & $0.5931$ & $-0.02161$ & $0.04190$ & $-2.968 \times 10^{-3}$ \\
 \hline
 \hline
\end{tabular}
\end{table*}

\begin{table*}
\centering
\caption{
I-Love-Q and C-Love, relations fit parameters for the constrained EoS data using the improved fitting relations found in Eq.~\ref{eq:ILQfitNew}.
This fitting relation, unlike previous versions, properly limits to the Newtonian case as $\Lambda \rightarrow \infty$.
}\label{tab:ILQfitNew}
\begin{tabular}{ c  c  | c c c c c c c c} 
 \hline
 \hline
 $y$ & $x$ & $\alpha$ & $K_{yx}$ & $a_1$ & $a_2$ & $a_3$ & $b_1$ & $b_2$ & $b_3$ \\
 \hline
 $\bar{I}$ & $\Lambda$ & $2/5$ & $0.5313$ & $1.287$ & $0.09888$ & $-2.300$ & $-1.347$ & $0.3857$ & $-0.02870$\\
 $\bar{Q}$ & $\Lambda$ & $1/5$ & $3.555$ & $-2.122$ & $2.72$ & $-1.491$ & $0.8644$ & $-0.1428$ & $-1.397$\\
 $\bar{I}$ & $\bar{Q}$ & $2$ & $0.008921$ & $10.59$ & $-37.46$ & $43.18$ & $-2.361$ & $1.967$ & $-0.5678$\\
 $C$ & $\Lambda$ & $-1/5$ & $0.2496$ & $-919.6$ & $330.3$ & $-857.2$ & $-383.5$ & $192.5$ & $-811.1$\\
\hline
\hline
\end{tabular}
\end{table*}

Figure~\ref{fig:ILQ} shows the improved I-Love-Q relations between the dimensionless moment of inertia $\bar{I} \equiv I/M^3$, the dimensionless quadrupole moment $\bar{Q} \equiv Q/M^3$, and the dimensionless tidal deformability $\Lambda$ for both the constrained and unconstrained sets of EoSs.
Here, each category of EoS are fit individually using the curve found in Eq.~\ref{eq:ILQfitNew}.
We observe that the constrained EoSs show considerable improvement from the unconstrained EoSs.
Additionally, we note a large improvement from the results of previous works~\cite{Yagi:ILQ}.
This is validated in Tab.~\ref{tab:maxVar}, where the maximal EoS variation from each fit is tabulated; comparing the results of previous works to the unconstrained and constrained sets of EoSs.
Observe how the constrained EoSs outperform both other cases by a considerable amount for each I-Love-Q relation.

\begin{figure*}
\begin{center} 
\includegraphics[width=.32\textwidth]{IL.pdf}
\includegraphics[width=.32\textwidth]{QL.pdf}
\includegraphics[width=.32\textwidth]{IQ.pdf}
\end{center}
\caption{
Individual I-Love-Q relations $\bar{I}-\Lambda$ (left), $\bar{Q}-\Lambda$ (center), and $\bar{I}-\bar{Q}$ (right), shown for both the constrained EoSs (solid green) and unconstrained EoSs (dotted maroon).
In these figures, the black dashed lines corresponds to the fits given by Eq.~\ref{eq:ILQfitNew}.
Observe how the fractional difference from the fits, shown in the bottom panels, is greatly suppressed for the constrained case, compared to both the unconstrained case, and results from previous works~\cite{Yagi:ILQ}.
The maximal EoS variation from the fits for the unconstrained and constrained sets of EoSs are compared in Tab.~\ref{tab:maxVar}.
Additionally shown in this figure is the fractional difference from the nuclear matter fits for the 10 hybrid star EoSs (dashed green).
}
\label{fig:ILQ}
\end{figure*}

\begin{table}
\centering
\caption{
Comparison between the EoS-insensitive relations (I-Love-Q, C-Love, R-Love, and binary Love) maximal EoS variability for the results of previous works~\cite{Yagi:ILQ,Yagi:binLove}, and the unconstrained and constrained sets of EoSs analyzed here. 
The maximum EoS variation, given by the largest fractional difference (the absolute difference is also shown in parenthesis for the C-Love, R-Love, and binary Love fits) from the fits as shown in Figs.~\ref{fig:ILQ},~\ref{fig:clove},~\ref{fig:rlove} and~\ref{fig:binLove}, observes a considerable improvement for the constrained set of EoSs, compared to the unconstrained set as well as that found in previous works.
Additionally, note how the maximal EoS variation for the unconstrained set of EoSs is slightly larger than that found in Refs.~\cite{Yagi:ILQ,Yagi:binLove} - due to the effects of large random sampling taking into account more sources of uncertainty.
}\label{tab:maxVar}
\begin{tabular}{ c  || c c c } 
 \hline
 \hline
 \textbf{EoS-insensitive} & \multicolumn{3}{c}{\textbf{Maximal EoS Variability}} \\
 \cline{2-4}
 \textbf{Relation} & \multicolumn{1}{c|}{\emph{Previous}} & \multicolumn{1}{c|}{\emph{Unconstrained}} & \emph{Constrained}\\
 \hline
 $\bar{I}-\Lambda$ &  $0.0059$ & $0.007673$ & $0.003138$\\
 $\bar{Q}-\Lambda$ & $0.010$ & $0.01262$ & $0.004661$\\
 $\bar{I}-\bar{Q}$ & $0.012$ & $0.01490$ & $0.005716$\\
 \hline
 \noalign{\smallskip}
\noalign{\smallskip}
 \noalign{\smallskip}
\noalign{\smallskip}
\hline
 \hline
 \textbf{EoS-insensitive} & \multicolumn{3}{c}{\textbf{Maximal EoS Variability}} \\
 \cline{2-4}
 \textbf{Relation} & \multicolumn{1}{c|}{\emph{Previous}} & \multicolumn{1}{c|}{\emph{Unconstrained}} & \emph{Constrained}\\
 \hline
 \multirow{2}{*}{$C-\Lambda$} & $0.065$ & $0.07198$ & $0.02169$\\
 & (--) & ($0.01837$) & ($0.006582$)\\
  \hline
 \multirow{2}{*}{$R-\Lambda$} & -- & $0.05649$ & $0.02172$\\
 & (--) & ($881.4 \text{ m}$) & ($358.9 \text{ m}$) \\
 \hline
 $\Lambda_a-\Lambda_s$ & $\sim0.50$ & $0.5724$ & $0.2140$\\
 $q=0.90$ & (--) & ($190.1$) & ($37.23$) \\
 \cline{1-1}
 $\Lambda_a-\Lambda_s$ & $\sim0.20$ & $0.2447$ & $0.08284$\\
  $q=0.7$ & (--) & ($320.4$) & ($51.62$) \\
  \cline{1-1}
 $\Lambda_a-\Lambda_s$ & $\sim0.025$ & $0.03841$ & $0.01823$\\
  $q=0.50$ & (--) & ($244.1$) & ($28.96$) \\
  \cline{1-1}
\hline
\hline
\end{tabular}
\end{table}

These results indicate that EoS-insensitive relations can indeed be greatly improved upon by restricting to EoSs in agreeance with physical observations.
In addition, we confirm the validity of this method by noting that the unconstrained EoSs observe more EoS variability than that found in Ref.~\cite{Yagi:ILQ} due to the nature of large random sampling taking into account more sources of uncertainty than was previously studied.
On the other hand, the constrained set of EoSs show significant improvement from that found in previous works.
This improvement can be beneficial to many future analyses, such as improvements to the constraints found on the moment of inertia, tidal deformability, compactness, spin, and radius found in Ref.~\cite{Kumar:2019xgp}.

\subsubsection{Hybrid quark-hadron stars}
\label{sec:ilq-hyb}

In this section, we investigate the I-Love-Q universality of hybrid stars, and their compatibility with their nuclear matter counterparts.
We consider three different sets of data to be fit by Eq.~\ref{eq:ILQfitNew}:
\begin{enumerate}
\item the complete set of 100 constrained EoSs combined with the 10 hybrid star EoSs,
\item the complete set of 100 constrained EoSs alone,
\item the complete set of 10 hybrid star EoSs alone.
\end{enumerate}
Following this, we compute the fractional difference from the fits for all three cases for the 10 hybrid star EoSs.
The fractional differences for the second case (fit to only the constrained EoSs) is shown by the dashed green lines in Fig.~\ref{fig:ILQ} for example.
For comparison purposes, Tab.~\ref{tab:hybridCompare} displays the maximal EoS variation in the $I-\Lambda$ relation for both the constrained and hybrid star EoSs in each case.

\begin{table}
\centering
\caption{
Comparison between the $I-\Lambda$ maximal EoS variability (fractional difference from the fit) for the constrained EoSs and the hybrid EoSs for three different cases of fitting data to Eq.~\ref{eq:ILQfitNew}: (1) fit to the combined data of constrained and hybrid EoS, (2) fit to the constrained EoS data alone, and (3) fit to the hybrid EoS data alone.
Observe how for all 3 cases, the hybrid EoSs are only universal up to a minimum of $\sim1$\%, while in each case the constrained EoSs outperform the hybrid ones.
The second case is demonstrated in Fig.~\ref{fig:ILQ}.
}\label{tab:hybridCompare}
\begin{tabular}{ c  || c c } 
 \hline
 \hline
 \textbf{Fitting} & \multicolumn{2}{c}{\textbf{Maximal EoS Variability}} \\
 \cline{2-3}
 \textbf{Case} &  \multicolumn{1}{c|}{\emph{Constrained}} & \emph{Hybrid}\\
 \hline
 \emph{Combined} &  \multirow{2}{*}{$0.004407$} & \multirow{2}{*}{$0.01363$}\\
 \emph{(Case 1)} & &\\
 \cline{1-1}
 \emph{Constrained only} & \multirow{2}{*}{$0.003138$} & \multirow{2}{*}{$0.01736$}\\
  \emph{(Case 2)} & &\\
  \cline{1-1}
 \emph{Hybrid only} & \multirow{2}{*}{$0.008400$} & \multirow{2}{*}{$0.01017$}\\
  \emph{(Case 3)} & &\\
  \cline{1-1}
\hline
\hline
\end{tabular}
\end{table}

Observe how the hybrid star EoSs obey the I-Love-Q relations in each case by up to $\sim1.7$\%, slightly higher than that found for nuclear matter EoSs in previous works~\cite{Yagi:ILQ}.
Further, we observe that the universality can not be improved by much through the introduction of new fits, only bringing the maximal EoS variation down to $\sim1$\% for the fits constructed with only hybrid star EoSs.
Concluding, we claim that hybrid star EoSs \emph{do} obey the traditional nuclear I-Love-Q relations computed through nuclear EoS data, with a slight decrease in universality to $\sim1.7$\%.

%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%% C-LOVE RELATIONS %%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{C-Love Relations}
\label{sec:clove}

In this section we consider approximately EoS-insensitive C-Love relations between the NS compactness $C \equiv M/R$ and the unit-less tidal deformability $\Lambda$, as discussed by Yagi and Yunes in ref.~\cite{Yagi:binLove}.
Additionally in this section, we compute the relationship between the NS radius and tidal deformability from the C-Love relations.
This allows one to compare the approximate uncertainty on the NS radius computed from the C-Love relations for both the constrained and unconstrained sets of EoSs. 
Similar to the previous section, we consider two classes of stars in this section: nuclear matter EoSs and hybrid quark-hadron star EoSs.

\subsubsection{Nuclear matter stars}
\label{sec:clove-nuc}

Following Ref.~\cite{Yagi:binLove}, we begin by fitting the data for each set of EoSs to the simple curve:
\begin{equation}
C = \sum^2_{k=0} a_k (\ln{\Lambda})^k.
\end{equation}
Doing so yields $a_0 = 0.3617$, $a_1 = -0.03548$, and $a_2 = 0.0006194$, similar to that found in Ref.~\cite{Yagi:binLove}.
Additionally we offer an improvement to the fits as was done in Sec.~\ref{sec:ilq}, by considering the fitting function found in Eq.~\ref{eq:ILQfitNew}, with the controlling factor given by:
\begin{equation}
C^N=K_{C\Lambda}\Lambda^{-1/5}.\label{eq:cloveFit}
\end{equation}
The new fitting coefficients are displayed in Tab.~\ref{tab:ILQfitNew}, alongside the previously found I-Love-Q relations.

Figure~\ref{fig:clove} presents the C-Love relations for both the constrained and unconstrained sets of EoSs, along with the corresponding absolute differences from the fits. 
Observe how the constrained EoSs show a large suppression in EoS variability compared to the unconstrained EoSs, as well as to that found in previous works~\cite{Yagi:binLove}.
The maximal EoS variation is compared between the above three cases in Tab.~\ref{tab:maxVar}.
\begin{figure}
\begin{center} 
\includegraphics[width=\columnwidth]{CL.eps}
\end{center}
\caption{
Similar to Fig.~\ref{fig:ILQ} but for the C-Love relations, and the absolute differences (rather than fractional) from the fit.
In this figure, we similarly construct fits to  Eq.~\ref{eq:ILQfitNew}, properly limiting to the Newtonian case $C \sim \Lambda^{-1/5}$.
Observe how, as before, the absolute difference from the fits is greatly suppressed for the constrained case, compared to both the unconstrained case, and results from previous works~\cite{Yagi:binLove}.
The maximal EoS variation from the fits for the unconstrained and constrained sets of EoSs are compared in Tab.~\ref{tab:maxVar} for both fractional and absolute differences.
Additionally shown in this figure is the hybrid star C-Love relations along with their absolute differences from the constrained EoS fits (dashed green).
}
\label{fig:clove}
\end{figure}

In addition, Fig.~\ref{fig:rlove} displays the relationship between the NS radius and tidal deformability for the constrained and unconstrained sets of EoSs.
This relationship is computed directly from the C-Love relations discussed above as $R(\Lambda)=m/C(\Lambda)$ for observed mass $m=1.48\text{ M}_{\odot}$ from GW170817.
We also present the uncertainty on a NS radius measurement using the C-Love relations in the bottom panel of Fig.~\ref{fig:rlove}, defined as the absolute difference from the fits.
This indicates that one can reliably measure the NS radius within $\sim 180 \text{ m} -350$ m, using the C-Love relations computed using the constrained set of EoSs.
The uncertainty quoted here is comparable to the $140\text{ m}$ error found in Ref.~\cite{Gamba:2019kwu}, which investigated the uncertainty owed to the choice of the SLy EoS model for the crust.
The radius uncertainty for the unconstrained EoSs of $\sim1\text{ km}$ however, would dominate the corresponding error budget on stellar radius compared to the crust uncertainties.
We further observe a large suppression in the radius measurement accuracy from the unconstrained set of EoSs to the constrained set, tabulated in Tab.~\ref{tab:maxVar} as the maximal EoS measurement uncertainty.

\begin{figure}
\begin{center} 
\includegraphics[width=\columnwidth]{RL.eps}
\end{center}
\caption{
Similar to Fig.~\ref{fig:clove} but for the R-Love relationship between the NS radius and tidal deformability.
This relationship is computed directly from the C-Love relations, and estimates the uncertainty on the NS radius as computed through the C-Love relations.
Observe how the absolute difference from the fit is reduced significantly between the unconstrained and constrained sets of EoSs, shown by the comparisons in Tab.~\ref{tab:maxVar}.
}
\label{fig:rlove}
\end{figure}

The results for the C-Love and R-Love relations, similar to that found in Sec.~\ref{sec:ilq}, shows considerable improvement in using the constrained EoSs over any unconstrained set.
We also note that the unconstrained EoSs display slightly larger EoS variability than found in Ref.~\cite{Yagi:binLove}, due to the large random sampling as discussed previously in Sec.~\ref{sec:ilq}.

\subsubsection{Hybrid quark-hadron stars}
\label{sec:clove-hyb}

In this section, we consider the effect of hybrid star EoSs on the C-Love relation.
Much like Sec.~\ref{sec:ilq-hyb}, we perform 3 separate fits (constrained and hybrid EoSs combined, constrained EoSs only, and hybrid EoSs only) and compare the maximal EoS variation for each set of EoSs.

Table~\ref{tab:hybridCompareClove} compares the maximal EoS variability for the constrained and hybrid star EoSs under the fitting conditions for each case described above. 
We observe very similar behavior to that described in Sec.~\ref{sec:ilq-hyb}, in which the maximal EoS variation for hybrid stars only fluctuate slightly ($\sim 4.5\% - 7\%$) for each case.
From this, we similarly conclude that hybrid stars do obey the C-Love relation derived with nuclear matter stars, with the caveat that the maximum universality increases to $\sim 7.1\%$.
This is exemplified in Fig.~\ref{fig:clove} by the dashed green curves.

\begin{table}
\centering
\caption{
Similar to Tab.~\ref{tab:hybridCompare} but for the C-Love relation.
Observe how for all 3 cases, the hybrid EoSs are only universal up to a minimum of $\sim1$\% (fractional difference from the fit), while in each case the constrained EoSs outperform the hybrid ones.
The second case is demonstrated in Fig.~\ref{fig:clove}.
}\label{tab:hybridCompareClove}
\begin{tabular}{ c  || c c } 
 \hline
 \hline
 \textbf{Fitting} & \multicolumn{2}{c}{\textbf{Maximal EoS Variability}} \\
 \cline{2-3}
 \textbf{Case} &  \multicolumn{1}{c|}{\emph{Constrained}} & \emph{Hybrid}\\
 \hline
 \emph{Combined} &  \multirow{2}{*}{$0.03686$} & \multirow{2}{*}{$0.05533$}\\
 \emph{(Case 1)} & &\\
 \cline{1-1}
 \emph{Constrained only} & \multirow{2}{*}{$0.02169$} & \multirow{2}{*}{$0.07191$}\\
  \emph{(Case 2)} & &\\
  \cline{1-1}
 \emph{Hybrid only} & \multirow{2}{*}{$0.05788$} & \multirow{2}{*}{$0.04478$}\\
  \emph{(Case 3)} & &\\
  \cline{1-1}
\hline
\hline
\end{tabular}
\end{table}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%% BINARY LOVE RELATIONS %%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Binary love relations}
\label{sec:binary}

Next we consider improvements to the binary Love relations.
Similar to Sec.~\ref{sec:ilq}, we consider two distinct classes of NSs: nuclear matter EoSs and hybrid quark-hadron star EoSs~\cite{Paschalidis2018,Alford:2017qgh,1971SvA....15..347S,Zdunik:2012dj,Alford:2013aca}.
Following Ref.~\cite{Yagi:binLove}, we begin by fitting the binary Love relations to the constrained and unconstrained sets of EoSs using the two-dimensional curve:
\begin{equation}\label{eq:binLovefit}
\Lambda_a=F_n(q) \frac{1+ \sum_{i=1}^3 \sum_{j=1}^2 b_{ij}q^j x^{i/5}}{1 + \sum_{i=1}^3 \sum_{j=1}^2 c_{ij}q^j x^{i/5}} \Lambda_s^{\alpha}.
\end{equation}
Here, q is the mass ratio $q \equiv m_1/m_2$ with $m_1 \leq m_2$, and $F_n(q)$ is the Newtonian-limiting control factor given by:
\begin{equation}\label{eq:control}
F_n(q) \equiv \frac{1-q^{10/(3-n)}}{1+q^{10/(3-n)}}.
\end{equation}
The updated fit parameters for the constrained set of EoSs can be found in Table~\ref{tab:binLovefit}.
Notice how, unlike the individual NS I-Love-Q relations, these relations additionally depend on the ratio of constituent masses in the binary system.

\begin{table*}
\centering
\caption{
Updated fit parameters for the binary Love relations, as given by the curve found in Eq.~\ref{eq:binLovefit}.
}\label{tab:binLovefit}
\addtolength{\tabcolsep}{1pt} 
\begin{tabular}{c  c  c  c  c  c  c  c c c c c c c} 
 \hline
 \hline
$n$ & $\alpha$ & $b_{11}$ & $b_{12}$ & $b_{21}$ & $b_{22}$ & $b_{31}$ & $b_{32}$ & $c_{11}$ & $c_{12}$ & $c_{21}$ & $c_{22}$ & $c_{31}$ & $c_{32}$\\
\hline
$0.743$ & $-1$ & $-14.40$ & $14.45$ & $31.36$ & $-32.25$ & $-22.44$ & $20.35$ & $-15.25$ & $15.37$ & $37.33$ & $-43.20$ & $-29.93$ & $35.18$\\
 \hline
 \hline
\end{tabular}
\addtolength{\tabcolsep}{-1pt}
\end{table*}

Figure~\ref{fig:binLove} shows the improved binary Love relation for 3 different mass ratios: $q=0.9$, $q=0.75$, and $q=0.5$ for both the constrained and unconstrained sets of EoSs.
Observe how once again, the constrained set of EoSs show a considerable improvement upon both the unconstrained set of EoSs, as well as the results found in previous works~\cite{Yagi:binLove}.
Additionally, the unconstrained set of EoSs show similar, yet slightly larger EoS variation than was previously found in Ref.~\cite{Yagi:binLove} due to the random sampling taken into account in this analysis.
Once again, the maximum EoS variability (in terms of both absolute and fractional differences from the fit) is tabulated in Tab.~\ref{tab:maxVar} for each value of mass ratio considered here.
We take note that while the fractional errors increase rapidly with increasing mass ratio $q$, the absolute errors remain consistent, with a subtle maximum at $q=0.75$. 
This can be explained by the control factor $F_n(q)$ given by Eq.~\ref{eq:control}, which similarly observes a peak for central mass ratio values of $q\approx0.75$.
We additionally make two observations about the effective ranges of $\Lambda_s$ seen in Fig.~\ref{fig:binLove}: (i) as the constituent masses in the binary system become further separated, the range of $\Lambda_s$ values shrinks; and (ii) the constrained set of EoSs typically display smaller values of $\Lambda_s$ than the unconstrained set do.
The former observation can be explained by the following reasoning: as the masses further separate, there is a limited range of tidal deformabilities $\Lambda_{1,2}$ in the M-L relationships which satisfy the mass ratio constraint.
From this we suggest the reader take note that as the mass ratio decreases, the allowable parameter space for the binary Love relations shrinks.
The latter observation originates from the EoS generation process: the unconstrained EoSs are randomly generated -- resulting in various EoSs which are both stiffer and softer than those constrained by GW170817.
This results in the unconstrained EoSs containing a larger range of tidal deformabilities $\Lambda$ (both above and below) than the constrained ones, which ultimately gives \emph{larger} values of $\Lambda_s=\frac{1}{2}(\Lambda_1+\Lambda_2)$.

\begin{figure}
\begin{center} 
\includegraphics[width=\linewidth]{binLoveAbsolute.eps}
\end{center}
\caption{
Binary Love relations shown for the constrained EoSs (dotted maroon) and unconstrained EoSs (solid green) for various values of mass ratio: $q=0.9$, $q=0.75$, and $q=0.50$.
In this figure, the top panel displays the EoS-insensitive relations with fits given by Eq.~\ref{eq:binLovefit} shown by dashed black lines (note there are 2 fits for each mass ratio, corresponding to the constrained and unconstrained sets of EoSs), while the bottom 3 panels correspond to the absolute EoS variation for each mass ratio.
Observe how the constrained set of EoSs show a reduction in EoS variation for both the unconstrained set, and from the results shown in Ref.~\cite{Yagi:binLove}.
The maximal EoS variability (shown as both the fractional and absolute differences) for each case is tabulated and compared in Tab.~\ref{tab:maxVar}.
Additionally shown are the binary Love relations for the 10 hybrid star EoSs (dashed bright green curves).
Observe the large deviations from the fit as the mass ratio $q$ increases, returning near the fit before the transitional pressure at large values of $\Lambda_s$.
}
\label{fig:binLove}
\end{figure}

Additionally shown in Fig.~\ref{fig:binLove} are the binary Love relations for the 10 hybrid star EoSs discussed in previous sections.
As described in Sec.~\ref{sec:theory}, transitional quark-hadron matter stars undergo first-order phase transitions at pressure $P_{\text{tr}}$, where the hadronic branch departs into a quark-matter branch at the corresponding transitional mass $M_{\text{tr}}$.
These transitions result in two distinct types of NSs, based on their observed mass: (i) massive ($M \geq M_{\text{tr}}$) hybrid stars which have quark-matter cores and nuclear matter crusts (we denote these as ``HS"); and (ii) low-mass ($M \leq M_{\text{tr}}$) hadronic stars with no internal transition to quark matter (we denote these as ``NS").
Contrary to the single-star relations (I-Love-Q, C-Love, and R-Love) where the hybrid stars moderately agree with their nuclear matter counterparts, the binary-star relations show large disagreement, which can be seen in Fig.~\ref{fig:binLove}.
Observe how the agreement with the fit is quite good at low pressures (large $\Lambda_s$) where the the binaries consist of NS/NS combinations (here, both stars' EoSs consist of the same class of nuclear matter as the constrained EoSs).
Once the critical pressure is reached, one or both stars transition into the hybrid branches, resulting in large reductions in tidal deformability as shown in the right panel of Fig.~\ref{fig:eos}.
When one star lies on the hadronic-matter branch while the other is on the quark-matter branch, 
the difference between tidal deformabilities becomes larger than expected for pure hadronic-matter stars.
This results in large deviations in the sums and differences of tidal deformabilities $(\Lambda_1 \pm \Lambda_2)$, disrupting the overall universality and generating the large ``bump" in the binary love relations for hybrid stars\footnote{This discrepancy is not present in the I-Love-Q and C-Love relations due to their single-star nature.} seen in Fig~\ref{fig:binLove}.

Due to the inconsistencies in the sums and differences of tidal deformabilities for binary hybrid star systems, we conclude that hybrid stars do \emph{not} agree with the traditional nuclear matter binary Love relations.
A number of attempts were made to reconcile these differences, such as performing separate ``HS" binary Love relations fitted directly to the hybrid star branches of the EoSs.
This analysis quickly left the scope of the paper due to the large number of parameters required to control the varying transitional pressures through different EoSs.
Therefore, we leave an analysis on the hybrid star binary Love relations to be discussed in future works.
There, we anticipate the generation of large numbers of hybrid star EoSs with varying transition masses to be taken into account as an additional fitting parameter in a piece-wise fitting function designed to take into account the ``bumps" seen in the binary Love relations.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%% FUTURE OBSERVATIONS %%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Impact on future observations}
\label{sec:observations}

We have now shown that binary NS merger observations can help improve the EoS-insensitive relations - the question is: is it worth it?
Current interferometer sensitivities are not yet small enough to accurately constrain the dominant tidal parameter $\tilde{\Lambda}$.
For example, GW170817 was detected by LIGO observing run 2 (``O2")~\cite{aLIGO} and was able to constrain $\tilde{\Lambda}$ to a $90\%$ confidence interval of 325 (or $\sigma_{\tilde{\Lambda}} \approx 198$) centered at $\mu_{\tilde{\Lambda}}=395$.
This corresponds to statistical uncertainties on the order of $\sim 82\%$, which dominates the error budget compared to the small systematic uncertainties picked up by EoS variation in the EoS-insensitive relations.
This implies that currently, the use of improved EoS-insensitive relations will only make a negligible difference on the extraction of tidal parameters.

In Sec.~\ref{sec:marginalization}, we first estimate the systematic uncertainties introduced by using the improved binary Love relations derived earlier.
In Sec.~\ref{sec:futureObservations}, we estimate the statistical uncertainties on the parameter extraction of tidal parameters, and compare them to the above-mentioned systematic uncertainties.
This is repeated for 5 future detectors, where multiple event detections become important.
Because future events occur with chirp tidal parameter $\tilde\Lambda$ which are unknown at this time, we can not accurately combine the uncertainties on $\tilde\Lambda$ for multiple events.
To remedy this, we re-parameterize the waveform and instead consider the $\lambda_0$ and $\lambda_1$ tidal coefficients produced by Taylor expanding the dimensionless tidal deformability $\Lambda$ about the ``canonical" reference mass of $m_0=1.4\text{ M}_{\odot}$~\cite{delPozzo:TaylorTidal,Yagi:binLove}:
\begin{equation}
\Lambda \approx \lambda_0 + \lambda_1 (1-\frac{m}{m_0}),
\end{equation}
where $\lambda_0$ and $\lambda_1$ are the tidal deformability and its slope at $1.4 \text{ M}_{\odot}$\footnote{$\lambda_0$ is commonly known as $\Lambda_{1.4}$.}. 
The new tidal parameters $\lambda_0$ and $\lambda_1$ do not depend on mass, thus are identical for every future binary NS merger, and may be combined in uncertainty.
Under this reasoning, we consider the new tidal parameters $\lambda_0$ and $\lambda_1$ for the remainder of the investigation.

\subsection{Marginalization over Uncertainties}\label{sec:marginalization}

In this section, we investigate the residuals on $\Lambda_a(\Lambda_s,q)$ in order to marginalize over the intrinsic uncertainty found in the binary Love EoS-insensitive relations.
In particular, we restrict our focus to only the constrained set of EoSs.
Residuals in $\Lambda_a$ are computed as $\Lambda_a^{\text{relation}}(\Lambda_s,q)-\Lambda_a^{\text{true}}$, where $\Lambda_a^{\text{true}}$ corresponds to the true value predicted by the sample EoSs, and $\Lambda_a^{\text{relation}}(\Lambda_s,q)$ corresponds to the value obtained through the EoS-insensitive relations found in Sec.~\ref{sec:binary}.
We follow this up by converting the $\Lambda_a$ residuals into $\lambda_0$, allowing us to approximate the overall systematic uncertainties on $\lambda_0$ from using the improved binary Love relations.

Following Ref.~\cite{Katerina:residuals}, we assume the residuals on $\Lambda_a$ observe a Gaussian distribution with mean and standard deviation given by:
\begin{align}
\mu_{\Lambda_a}(\Lambda_s,q) &=\frac{\mu_{\Lambda_s}(\Lambda_s)+\mu_{q}(q)}{2},\\ 
\sigma_{\Lambda_a} &=\sqrt{\sigma_{\Lambda_s}^2(\Lambda_s) + \sigma_{q}^2(q)}. 
\end{align}
Similar to Ref.~\cite{Katerina:residuals}, we fit the individual components to the following power functions:
\begin{align}
\mu_{\Lambda_s}(x) &= \mu_1 x + \mu_2, \label{eq:margFit1}\\ 
\mu_{q}(x) &= \mu_3 x^2 + \mu_4 x + \mu_5, \label{eq:margFit2}\\ 
\sigma_{\Lambda_s}(x) &= \sigma_1 x^{5/2} + \sigma_2 x^{3/2} + \sigma_3 x \sigma_4 x^{1/2} + \sigma_5, \label{eq:margFit3}\\ 
\sigma_{q}(x) &= \sigma_6 x^3 + \sigma_7 x^2 + \sigma_8 x + \sigma_9. \label{eq:margFit4}
\end{align}
The fitting parameters $\mu_i$ and $\sigma_i$ are tabulated in Table~\ref{tab:marginalized}.

\begin{table}
\centering
\caption{
Coefficients to the fits given by Eqs.~(\ref{eq:margFit1})-(\ref{eq:margFit4}) for the relative error on $\Lambda_a$ in the improved binary Love EoS-insensitive relations presented in this paper.
}\label{tab:marginalized}
\addtolength{\tabcolsep}{1pt} 
\begin{tabular}{ c | c || c | c}
\hline 
\noalign{\smallskip}
$\mu_1$ & $3.509 \times 10^{-3}$ & $\sigma_1$ & $-2.074 \times 10^{-7}$\\
$\mu_2$ & $9.351 \times 10^{-1}$ & $\sigma_2$ & $-1.492 \times 10^{-3}$\\
$\mu_3$ & $-18.07$ & $\sigma_3$ & $-4.891 \times 10^{-2}$\\
$\mu_4$ & $27.56$ & $\sigma_4$ & $8.207 \times 10^{-1}$\\
$\mu_5$ & $-10.10$ & $\sigma_5$ & $-1.308$\\
 &  & $\sigma_6$ & $-63.76$\\
 &  & $\sigma_7$ & $11.14$\\
 &  & $\sigma_8$ & $75.25$\\
 &  & $\sigma_8$ & $-23.69$\\
 \noalign{\smallskip}
 \hline
\end{tabular}
\addtolength{\tabcolsep}{-1pt}
\end{table}

Now that we have obtained the residuals on $\Lambda_a$, we must convert them into the new tidal parameter $\lambda_0$, in order to compute and easily compare the systematic uncertainties to the statistical uncertainties on $\lambda_0$, found in Sec.~\ref{sec:futureObservations}.
Figure~\ref{fig:qLsResiduals} shows the standard deviations of these residuals (converted into $\lambda_0$) binned in both $q$ and $\Lambda_s$ as was done above. 
To estimate the overall systematic uncertainties on $\lambda_0$ accrued upon the use of the improved binary Love relations, we combine the uncertainties in both $q$ and $\Lambda_s$.
Figure~\ref{fig:residuals} displays the un-binned Gaussian distribution of $\lambda_0$ residuals for both the constrained, and unconstrained sets of EoSs for comparison.
Observe how the standard deviations $\sigma=9.764$ and $\sigma=78.28$ show a large decrease from the unconstrained to the constrained sets of EoSs.
Additionally, we find the 90th, 99th, and 100th percentiles on $\lambda_0$ to be $P_{90}=13.19$, $P_{99}=42.38$, and $P_{100}=111.32$ for the constrained EoSs.
Because the un-binned residuals seen in Fig.~\ref{fig:residuals} are dominated by the low-error regions of the parameter space ($\Lambda_s \rightarrow 0$, $\Lambda_s \rightarrow \infty$, $q \rightarrow 0$ and $q \rightarrow 1$) shown by Fig.~\ref{fig:qLsResiduals}, we consider the 90th percentile uncertainty (indicated by the dashed green line in Fig.~\ref{fig:qLsResiduals}) for the remainder of the analysis, in order to further take into account the high-error regions of the parameter space ($\Lambda_s \sim 2000$, $q\sim0.5$).
We take this value of $P_{90}=13.19$ to be the systematic measurement uncertainty introduced by using the improved binary Love relations on the parameter extraction of $\lambda_0$, indicated by the dashed indigo line in Fig.~\ref{fig:stackedFisher}.
Similarly, we derive the systematic measurement uncertainty on $\tilde\Lambda$ to be $12.01$, shown in Fig.~\ref{fig:singleFisherLt}.
Currently, the statistical uncertainty of $170.1$ on $\lambda_0$ from GW170817 dominates the error budget compared to the systematic uncertainty of $13.19$, as seen in Fig.~\ref{fig:stackedFisher}.
This means that, for the current ``O2" detector sensitivity, the use of improved binary Love relations will only make a negligible difference on the extraction of $\tilde\Lambda$.
However, future detectors (for example aLIGO, A\texttt{+}, Voyager, CE, and ET) are planned with large reductions in sensitivity - both decreasing the statistical uncertainties and allowing for larger binary NS merger detection rates; further reducing the statistical uncertainties.
In Sec.~\ref{sec:futureObservations}, we analyze this further and discuss when the systematic uncertainties from EoS-insensitive relations become comparable to the statistical uncertainties.

\begin{figure}
\begin{center} 
\includegraphics[width=\columnwidth]{qResiduals.eps}
\includegraphics[width=\columnwidth]{LsResiduals.eps}
\end{center}
\caption{$\lambda_0$ residuals binned in $q$ (top) and $\Lambda_s$ (bottom), highlighting the different error weights across the entire $(q,\Lambda_s)$ parameter space.
In this figure, the violet circles indicate the standard deviation of each bin in $q$-space ($\Lambda_s$-space), while the solid black lines represent the best fit given by Eqs. (\ref{eq:margFit3})-(\ref{eq:margFit4}).
Observe how the uncertainty is maximal for $q\sim0.5$ and $\Lambda_s\sim2000$, while it becomes minimal for both low and high values of $q$ and $\Lambda_s$.
Also shown in the figure is the 90th percentile of the un-binned residuals seen in Fig.~\ref{fig:residuals}, taken to be the overall systematic uncertainty introduced by using binary Love relations.
}
\label{fig:qLsResiduals}
\end{figure}

\begin{figure}
\begin{center} 
\includegraphics[width=\columnwidth]{residuals.pdf}
\end{center}
\caption{
Residuals on $\lambda_0$ computed as $\lambda_0^{\text{fit}}-\lambda_0^{\text{true}}$ for the binary Love relations modeled in Sec.~\ref{sec:binary} for both constrained and unconstrained sets of EoSs.
These residuals obey Gaussian distributions centered at $\mu=-0.1530$ and $\mu=0.2710$ with standard deviations of $\sigma=9.764$ and $\sigma=78.28$ for the constrained and unconstrained sets of EoSs, respectively.
These uncertainties correspond roughly to the systematic uncertainties introduced on the parameter extraction of $\lambda_0$ upon the use of binary Love relations.
However, to take into account the systematic uncertainties found in high-error regions of the parameter space, we instead set the systematic uncertainty to be the 90th percentile, $P_{90}=13.19$.
Observe that the systematic uncertainties from using the improved (constrained) binary Love relations are negligible compared to the statistical uncertainties accrued on parameter extraction from GW170817, found to be $\sigma_{\lambda_0}=170.1$.
}
\label{fig:residuals}
\end{figure}

\subsection{Future Observations}\label{sec:futureObservations}
In this section, we estimate the feasibility of utilizing the improved EoS-insensitive relations in future binary NS merger events.
This estimate is acquired through a simple Fisher analysis~\cite{Finn:Fisher,Cutler:Fisher} which approximates the accuracy with which one can extract the best-fit parameters $\theta^a$, given a prior template waveform.
For the remainder of the paper, we consider a template parameter vector consisting of:
\begin{equation}\label{eq:template}
\theta^a=(\ln{A},\phi_c,t_c,\ln{\mathcal{M}},\ln{\mathcal{\eta}},\chi_s,\chi_a,\lambda_0, \lambda_1),
\end{equation}
where $\eta \equiv m_1 m_2/M^2$ is the symmetric mass ratio with $m_{1,2}$ and $M$ being the individual and total masses, $\mathcal{M}=M \eta^{3/5}$ is the chirp mass, $A \equiv \frac{\mathcal{M}^{5/6}}{\sqrt{30}\pi^{2/3}D_L}$ is a normalized amplitude factor with $D_L$ being the luminosity distance, and $\chi_{s,a}=\frac{1}{2}(\chi_1\pm\chi_2)$ are the symmetric and antisymmetric total spins, where $\chi_{1,2}$ are the NSs individual spins. 
Following Refs.~\cite{Cutler:Fisher,Berti:Fisher,Poisson:Fisher}, this method relies on the crude assumption of Gaussian prior distributions\footnote{Typically, the more valid assumption is a uniform prior distribution; an improvement made in a more detailed Bayesian analysis.}.
The resulting posterior distribution is Gaussian with a root-mean-square given by:
\begin{equation}
\Delta \theta^a=\sqrt{\Big( \tilde{\Gamma}^{-1}\Big)^{aa}}.
\end{equation}
Here, the Fisher matrix $\tilde{\Gamma}$ is defined by:
\begin{equation}
\tilde{\Gamma}_{ab} \equiv \Big( \frac{\partial h}{\partial \theta^a} \Big| \frac{\partial h}{\partial \theta^a}\Big) + \frac{1}{\sigma_{\theta^a}^2} \delta_{ab}
\end{equation}
where $h$ is the waveform template, $\sigma_{\theta^a}$ are the parameters' prior root-mean-square estimate, and the inner product is defined by:
\begin{equation}
(a|b) \equiv 2 \int^{\infty}_0\frac{\tilde{a}^*\tilde{b}+\tilde{b}^*\tilde{a}}{S_n(f)}df.
\end{equation}

In this analysis, we consider the ``PhenomD" (IMRD) waveform template~\cite{PhenomDI,PhenomDII} modified by two different tidal corrections: the 6PN tidal correction shown in Ref.~\cite{Wade:tidalCorrections} (IMRD + 6PN), and the NRTidal correction shown in Ref.~\cite{Samajdar:NRTidal} (IMRD + NRTidal).
Comparing the two different tidal corrections to the IMRD waveform returns a rough estimate on the magnitude of waveform mismodeling systematics.
Lastly, we consider the spectral noise densities $S_n^A(f)$ for 6 different interferometer designs: $A \equiv ($O2~\cite{aLIGO}, aLIGO~\cite{aLIGO}, A\texttt{+}~\cite{Ap_Voyager_CE}, Voyager~\cite{Ap_Voyager_CE}, CE~\cite{ET}, ET-D~\cite{Ap_Voyager_CE}$)$ in order to compare the statistical uncertainties accrued on parameter extraction using future upgraded LIGO detectors as well as third generation detectors.

We begin by authenticating this approach by applying a Fisher analysis (with an IMRD + 6PN waveform injection) to GW170817, as observed by LIGO with O2 detector sensitivity~\cite{aLIGO}.
Because only 1 event was detected, we utilize $\tilde\Lambda$ and $\delta\tilde\Lambda$ as the tidal parameters for comparison purposes.
Further, we scale the luminosity distance such that the signal-to-noise-ratio ($SNR \equiv \rho$) is fixed to be $\rho=32.4$, as was found in GW170817.
We also assume low spin priors $|\chi| \leq 0.05$, as well as $\tilde{\Lambda} \leq 3000$ and $|\delta \tilde{\Lambda}| \leq 500$~\cite{Wade:LambdaPriors}.
The resulting posterior distribution on $\tilde{\Lambda}$ has a range of $\pm 276.99$ encompassing the $90\%$ credible levels.
We compare these results with the Bayesian analysis performed in Ref.~\cite{TheLIGOScientific:2017qsa,Abbott2018}, finding close agreement with the resulting $90\%$ credible region of $70 \leq \tilde{\Lambda} \leq 720$.
This confirms the approximate validity of this method, allowing a continuation of this analysis for future detectors.
\begin{figure}
\begin{center} 
\includegraphics[width=\columnwidth]{sensitivities.eps}
\end{center}
\caption{
Square root of of the spectral noise densities $\sqrt{S_n^A(f)}$ plotted for detectors ($A$): LIGO O2, aLIGO, A\texttt{+}, Voyager, CE, and ET-D as interpolated from publicly available data.
Spectral noise densities are plotted from $f_{\text{min}}=(23,10,10,7,1,1) \text{ Hz}$, respectively, to $f_{\text{max}}=1649 \text{ Hz}$.
Also shown is $2 \sqrt{f}$ multiplied by the amplitude of the PhenomD~\cite{PhenomDI,PhenomDII} waveform template.
}
\label{fig:sensitivities}
\end{figure}

Next, we consider events identical to GW170817 detected on future design sensitivity upgrades and detectors $A \equiv ($O2, aLIGO, A\texttt{+}, Voyager, CE, ET-D$)$, shown in Fig.~\ref{fig:sensitivities}.
Further, we consider the combined statistical uncertainties of $N_A$ events detected over 1 year on detector $A$, determined by integrating the local binary NS (BNS) merger rate over redshifts up to the horizon redshift of each detector.
Here we assume low spin priors $|\chi| \leq 0.05$, as well as $0 \leq \lambda_0 \leq 3207$ and $-4490 \leq \lambda_1 \leq 0$~\cite{delPozzo:TaylorTidal}\footnote{These are converted from the dimensional forms found in Ref.~\cite{delPozzo:TaylorTidal} into their corresponding dimensionless forms.}.
The process we utilize to compute single and combined statistical uncertainties for each detector sensitivity $S_n^A(f)$ is detailed in App.~\ref{app:stackingProcedure}.
From this, we determine if and when the statistical uncertainties associated with the parameter extraction of $\lambda_0$ drop below the systematic EoS variation errors from using binary Love relations.

The results for this process upon the injection of the IMRD + 6PN waveform are compiled in Table~\ref{tab:variances}, and shown graphically in Fig.~\ref{fig:stackedFisher} where $\sigma^A_{\text{GW170817}}$ and $\sigma^A_N$ are plotted as a function of $\rho^A_{\text{GW170817}}$ (the SNR as if GW170817 were detected on future detector $A$) for 5 future detector sensitivities.
For reference, Fig.~\ref{fig:singleFisherLt} also displays the estimated $\tilde\Lambda$ extraction efficiency for \emph{one} event on each detector; as these events can not be combined like was done in the case of $\lambda_0$.
Observe how this figure shows similar results on $\tilde\Lambda$ to that found on $\lambda_0$ for the case of one event - that is, future third-generation detectors ET and CE both observe statistical uncertainties on $\tilde\Lambda$ (and $\lambda_0$) dropping below the systematic binary Love errors for a single event.
This result is amplified when multiple events are combined as shown in Fig.~\ref{fig:stackedFisher}, further motivating the need to improve the systematic uncertainties in the binary Love relations for upcoming detectors Voyager, CE and ET.
We note that in Figs.~\ref{fig:stackedFisher} and~\ref{fig:singleFisherLt}, the single-event measurement uncertainties are slightly larger for CE than for ET despite having a larger SNR.
This relic originates from the high-frequency dependence of CE and ETs detector sensitivities: due to the 3-detector geometry of ET, it is able to more accurately measure the tidal effects emergent at high frequencies.
However, when computing the SNRs from $1 \text{ Hz} \leq f \leq 1649 \text{ Hz}$, CE dominates over ET and observes a larger SNR.
When instead setting the minimum frequency to $f_{\text{min}} \geq 68$ Hz however, the SNR for ET instead dominates that for CE, validating our hypothesis on this effect.

\begin{figure}
\begin{center} 
\includegraphics[width=\columnwidth]{singleFisherLt.eps}
\end{center}
\caption{
Estimated statistical uncertainties $\sigma^A_{\text{GW170817}}$ in the extraction of $\tilde\Lambda$ from GW170817 (O2) as if observed with future interferometers aLIGO, A\texttt{+}, Voyager, CE, and ET-D as a function of the signal-to-noise-ratio $\rho^A_{\text{GW170817}}$.
Also shown is the systematic uncertainty on $\tilde\Lambda$ derived using the same method used to compute the systematic uncertainty on $\lambda_0$.
This is similar to Fig.~\ref{fig:stackedFisher} where the extraction efficiency of $\lambda_0$ was plotted, however, in this case multiple events can \emph{not} be combined due to the variation in mass and therefore $\tilde\Lambda$ among events.
}
\label{fig:singleFisherLt}
\end{figure}

\begin{table*}
\centering
\caption{
Tabulated results for the estimated statistical uncertainties on the tidal deformability at $1.4\text{ M}_{\odot}$, $\lambda_0$, for the combined results of $N_A$ detections corresponding to the estimated upper, central, and lower limits of the binary NS merger detection rate $\mathcal{R}$.
This is repeated for 5 future detector sensitivities: aLIGO, A\texttt{+}, Voyager, CE, and ET.
Below, $\rho^A_{\text{GW170817}}$ and $\sigma^A_{\text{GW170817}}$ correspond to the approximate SNR and $\sigma_{\lambda_0}$ of GW170817 had it been observed by future interferometer $A$, and $\sigma^A_N$ corresponds to the overall uncertainty on $\lambda_0$ after $N_A$ detections on interferometer $A$.
It can be seen here that the statistical uncertainties on $\lambda_0$ become comparable with the systematics (set to be $P_{90}=13.19$) from using improved binary Love relations for future detectors Voyager, ET, and CE, indicating when it will become important to improve the EoS-insensitive relations.
Note that the approximation for $\sigma^A_N$ under the O2 detector sensitivity is left blank due to the singularity of the detected event.
These results are also shown graphically in Fig.~\ref{fig:stackedFisher}.
}\label{tab:variances}
\begin{tabular}{|c|@{\extracolsep{4pt}}c@{\extracolsep{0pt}}|c|@{\extracolsep{4pt}}C{1.7cm}@{\extracolsep{-2pt}}|C{1.7cm}|@{\extracolsep{-2pt}}C{1.7cm}@{\extracolsep{2pt}}|c|@{\extracolsep{0pt}}c@{\extracolsep{0pt}}|c|}
\cline{1-1}\cline{2-3}\cline{4-9}
    \multicolumn{1}{|c|}{\bfseries Detectors (A)} & \multicolumn{2}{|c|}{\bfseries GW170817} & \multicolumn{6}{|c|}{\bfseries Multiple events} \\
\cline{1-1}\cline{2-3}\cline{4-9}
\noalign{\smallskip}
\cline{2-3}\cline{4-6}\cline{7-9}
\multicolumn{1}{c}{} & \multicolumn{1}{|c|}{} & \multicolumn{1}{c|}{} & \multicolumn{3}{|c|}{} & \multicolumn{3}{|c|}{}
\\[-1em]
\multicolumn{1}{c}{}  &  \multicolumn{1}{|c|}{\multirow{2}{*}{$\rho^A_{\text{GW170817}}$}}  &  \multirow{ 2}{*}{$\sigma^A_{\text{GW170817}}$}  &  \multicolumn{3}{|c|}{$N_A$}  &  \multicolumn{3}{|c|}{$\sigma^A_N$}  \\
\cline{4-6}\cline{7-9}
\multicolumn{1}{c}{}  &  \multicolumn{1}{|c|}{}  &  \multicolumn{1}{c|}{}  &  \multicolumn{1}{|c|}{Low}  &  \multicolumn{1}{c|}{Central} &  \multicolumn{1}{c|}{High}  & \multicolumn{1}{|c|}{Low}  &  \multicolumn{1}{c|}{Central} &  \multicolumn{1}{c|}{High}\\
\cline{2-3}\cline{4-6}\cline{7-9}
\noalign{\smallskip}
\noalign{\smallskip}
\cline{1-1}\cline{2-3}\cline{4-6}\cline{7-9}
 O2  & \multicolumn{1}{|c|}{$32.40$}  & $170.13$ &  \multicolumn{1}{|c|}{--} & -- & -- & \multicolumn{1}{|c|}{--} & -- & --\\
\cline{1-1}\cline{2-3}\cline{4-6}\cline{7-9}
 aLIGO  & \multicolumn{1}{|c|}{$90.74$}  & $109.7$ &  \multicolumn{1}{|c|}{$20$} & $98$ & $304$ & \multicolumn{1}{|c|}{$182.8$} & $83.31$ & $46.95$\\
\cline{1-1}\cline{2-3}\cline{4-6}\cline{7-9}
 A\texttt{+}  & \multicolumn{1}{|c|}{$180.95$}  & $45.64$ &  \multicolumn{1}{|c|}{163} & 786 & 2,419 & \multicolumn{1}{|c|}{$58.94$} & $24.69$ & $13.66$\\
\cline{1-1}\cline{2-3}\cline{4-6}\cline{7-9}

\multicolumn{1}{|c|}{} & \multicolumn{1}{|c|}{} & \multicolumn{1}{c|}{} & \multicolumn{1}{|c|}{} & \multicolumn{1}{c|}{} & \multicolumn{1}{c|}{} & \multicolumn{1}{|c|}{} & \multicolumn{1}{c|}{} & \multicolumn{1}{c|}{}
\\[-1em]

 Voyager  & \multicolumn{1}{|c|}{$428.89$}  & $25.26$ &  \multicolumn{1}{|c|}{2,191} & 10,544 & 32,454 & \multicolumn{1}{|c|}{$20.97$} & $9.556$ & $5.323$\\
\cline{1-1}\cline{2-3}\cline{4-6}\cline{7-9}

\multicolumn{1}{|c|}{} & \multicolumn{1}{|c|}{} & \multicolumn{1}{c|}{} & \multicolumn{1}{|c|}{} & \multicolumn{1}{c|}{} & \multicolumn{1}{c|}{} & \multicolumn{1}{|c|}{} & \multicolumn{1}{c|}{} & \multicolumn{1}{c|}{}
\\[-1em]

 ET-D  & \multicolumn{1}{|c|}{$1398.03$}  & $6.871$ &  \multicolumn{1}{|c|}{71,536} & 344,270 & 1,059,638 & \multicolumn{1}{|c|}{$3.848$} & $1.732$ & $0.9621$\\
\cline{1-1}\cline{2-3}\cline{4-6}\cline{7-9}

\multicolumn{1}{|c|}{} & \multicolumn{1}{|c|}{} & \multicolumn{1}{c|}{} & \multicolumn{1}{|c|}{} & \multicolumn{1}{c|}{} & \multicolumn{1}{c|}{} & \multicolumn{1}{|c|}{} & \multicolumn{1}{c|}{} & \multicolumn{1}{c|}{}
\\[-1em]

 CE  & \multicolumn{1}{|c|}{$2807.45$}  & $7.664$ &  \multicolumn{1}{|c|}{295,200} & 1,420,650 & 4,372,653 & \multicolumn{1}{|c|}{$3.720$} & $1.668$ & $0.8979$\\
\cline{1-1}\cline{2-3}\cline{4-6}\cline{7-9}
\end{tabular}
\end{table*}

Next, we repeat the process upon the injection of the IMRD + NRTidal waveform in order to demonstrate the effect of waveform mismodeling.
The results are shown in Fig.~\ref{fig:stackedFisher} as maroon outlined circles (single detection uncertainties) and an outlined dashed maroon region (combined uncertainties from multiple detections).
Observe the noticeable difference between the extraction efficiencies for the two different tidal corrections - highlighting the need for further waveform modeling for future interferometers.

Concluding, we find that the Voyager, ET, and CE detectors all exhibit enough statistical uncertainty reduction that the systematics on $\lambda_0$ and $\tilde\Lambda$ introduced from using improved EoS-insensitive relations (found to be $13.19$ and $12.01$, respectively) no longer become negligible in the error budget.
This implies that the use and further improvement of EoS-insensitive relations is not only justified, but necessary, for future binary NS merger detections.
Additionally, we find that the effect of waveform mismodeling is indeed important for future observations as the statistical uncertainties continue to decrease.
In Appendix~\ref{app:measurement}, we additionally investigate the dependence of the measurement accuracy on $\tilde\Lambda$ on the binary NS mass ratio $q$ for fixed chirp mass $\mathcal{M}=1.88\text{ M}_{\odot}$.

%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%
%%%% DISCUSSION %%%%
%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%

\section{Conclusion and Discussion}
\label{sec:conclusion}

The recent GW observation of binary NS merger GW170817 placed constraints on the supranuclear matter EoS for NSs.
We take advantage of this by generating a restricted set of spectral EoSs which agree with this observation in order to reduce the uncertainties upon the extraction of tidal parameters from future GW events.
Important previous work by Yagi and Yunes~\cite{Yagi:ILQ,Yagi:binLove} found EoS-insensitive relations between symmetric and antisymmetric combinations of NS tidal deformabilities, which aid in the extraction of said tidal parameters.
We find an improvement upon these EoS-insensitive relations by a factor of $\sim 59$\% for stars with mass ratios of $0.75$ by restricting to the EoS posterior samples from GW170817s 90\% credible region on pressure as a function of density.
Similarly, we find an increase in C-Love and I-Love-Q universality by factors of $\sim 76$\% and $\sim 50$\%, respectively.
These improvements on the I-Love-Q and C-Love relations can be very beneficial in many important analyses, including that done in Ref.~\cite{Kumar:2019xgp}, where constraints on the moment of inertia, tidal deformability, compactness, spin, and radius were found using such relations.
We also show the relationship between the NS radius and its tidal deformability, computed directly from the C-Love relations.
We find this relation to display a maximum uncertainty of $358.9\text{ m}$ from the fit discussed in Sec.~\ref{sec:clove}.
Additionally, we find that hybrid quark-hadron star EoSs \emph{do} obey the I-Love-Q and C-Love relations, albeit with slightly higher maximum EoS variabilities as displayed in Tabs.~\ref{tab:hybridCompare} and ~\ref{tab:hybridCompareClove}.
We additionally find that binaries consisting of one massive hybrid quark-hadron star, and one small-mass hadron star do not obey the binary Love relations -- explained by the large difference in $\tilde{\Lambda}$ displayed between the constituent stars.
From this we state that the hybrid star EoSs do \emph{not} obey the binary Love relations.
Future work currently in progress will attempt to remedy this through a series of new piecewise hybrid star binary Love relations.

In the second half of this paper, we analyze the impact of this improvement for future binary NS merger detections.
Currently detectors are not yet sensitive enough to accurately constrain $\tilde{\Lambda}$ for the improved EoS-insensitive relations to make a difference.
We find that future interferometers Voyager, CE, and ET-D all experience a reduction in the statistical uncertainties upon the extraction of $\tilde{\Lambda}$ enough to become comparable to the systematic uncertainties injected through the use of improved EoS-insensitive relations, as depicted in Fig.~\ref{fig:stackedFisher} and Table~\ref{tab:variances}.
Additionally, we considered the effect of waveform mismodeling by comparing Fisher analyses using two different tidal corrections to the PhenomD waveform: 6PN and NRTidal.
Doing so revealed a noticeable difference between the extraction uncertainties derived under injection of each waveform - indicating the need for further waveform modeling.
Finally, in Appendix~\ref{app:measurement} we investigate the effect of NS binary mass ratio q on the measurement accuracy of $\tilde\Lambda$.
We find that the second generation interferometers lose accuracy as one increases the mass ratio, while the third generation detectors observe a large increase in accuracy due to the correlations between $\tilde\Lambda$ and $\delta\tilde\Lambda$.
This indicates that the use and further improvement of EoS-insensitive relations is justified for future GW observations of binary NS merger events.

Future work on this subject entails an investigation into the improvement of alternative EoS-insensitive relations, such as the multipole Love relations between various $l$-th order electric, magnetic, and shape tidal deformabilities as discussed in Ref.~\cite{Yagi:Multipole}.
Additionally, Lackey \emph{et al}. presented surrogate models of non-spinning effective-one-body waveform models with the the use of universal relations in Refs.~\cite{Lackey:Surrogate, Lackey:EOB}.
By reducing the number of required waveform model parameters, surrogate models aid in the extraction of NS observables from GW detections.
The improvement in the multipole love relations can then be used to increase the accuracy of such surrogate models.
Additionally, we plan on a more comprehensive analysis into the intricacies of new hybrid star binary Love relations, as discussed in Sec.~\ref{sec:binary}.
Work along this direction is currently in progress.

%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%% ACKNOWLEDGEMENTS %%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%

\section*{Acknowledgments}\label{acknowledgments}
K.Y. acknowledge networking support by the COST Action GWverse CA16104.
N.Y. acknowledges support from  NSF grant PHY-1759615 and NASA grants NNX16AB98G and 80NSSC17M0041.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%% APPENDIX A: STACKED EVENTS PROCESS %%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\appendix

\section{Computation of statistical uncertainties}\label{app:stackingProcedure}

In this appendix, we detail the process used to compute the statistical uncertainties on the extraction of $\lambda_0$ from the gravitational waveform.
This is accomplished with a simple Fisher analysis described in Sec.~\ref{sec:futureObservations}, where we first consider single-events identical to GW170817 as if they were detected on future detectors $A \equiv ($ O2~\cite{aLIGO}, aLIGO~\cite{aLIGO}, A\texttt{+}~\cite{Ap_Voyager_CE}, Voyager~\cite{Ap_Voyager_CE}, ET~\cite{ET}, CE~\cite{Ap_Voyager_CE}$)$.
We conclude by simulating a population of $N_A$ events for each interferometer, approximating the number of events detected on interferometer $A$ over an observing period of one year.
This is used to combine the statistical uncertainties, resulting in an approximation on the overall measurement accuracy of $\lambda_0$, which is compared to the systematic uncertainties computed in Sec.~\ref{sec:marginalization}.
The process used to achieve this is outlined below:

\begin{enumerate}
\item[(i)] Perform a Fisher analysis as outlined in Sec.~\ref{sec:futureObservations} using detector sensitivity $S_n^A(f)$, while restricting the luminosity distance $D_L$ such that an SNR of $\rho^{\text{O2}}_{\text{GW170817}}=32.4$ would be achieved on O2 sensitivity $S_n^{\text{O2}}(f)$.
Here we assume low spin priors $|\chi_{1,2}| \leq 0.05$, as well as $0 \leq \lambda_0 \leq 3207$ and $-4490 \leq \lambda_1 \leq 0$~\cite{delPozzo:TaylorTidal} (these are converted from their corresponding dimensional forms).
This results in an SNR $\rho^A_{\text{GW170817}}$ and a single-event statistical uncertainty $\sigma_\text{GW170817}^A$ accrued in the extraction of $\lambda_0$ on detector $A$.

\item[(ii)] Generate a population of $N_A$ events corresponding to the expected binary NS merger detection rate for detector $A$, following the probability distribution~\cite{Shutz:SNR,Chen:SNR}:
\begin{equation}\label{eq:population}
f(\rho)=3 \rho_{\text{th}}^3/\rho^4
\end{equation}
with a network SNR threshold of $\rho_{\text{th}}=8$.
The number of events $N_A$ is calculated by taking into account the BNS merger rate history throughout all redshift values within detector As' horizon redshift $z_h$, as shown by Eq. (10) of Ref.~\cite{Cutler:BNSmerger}:
\begin{equation}
N_A=\Delta \tau_0 \int\limits^{z_{h}}_0 4 \pi \lbrack  a_0r_1(z)\rbrack^2 \mathcal{R} r(z) \frac{d \tau}{dz} dz.
\end{equation}
Here, $a_0r_1(z)$, $\frac{d\tau}{dz}$, and $r(z)$ for our chosen cosmology are given by:
\begin{equation}
a_0r_1(z) = \frac{1}{H_0}\int\limits^z_0 \frac{dz'}{\sqrt{(1-\Omega_{\Lambda})(1+z')^3+\Omega_{\Lambda}}},
\end{equation}
\begin{equation}
\frac{d\tau}{dz} = \frac{1}{H_0} \frac{1}{1+z}\frac{1}{\sqrt{(1-\Omega_{\Lambda})(1+z')^3+\Omega_{\Lambda}}},
\end{equation}
\begin{equation}
r(z) = \left\{
\begin{array}{ll}
      1+2z & z \leq 1 \\
      \frac{3}{4}(5-z) & 1\leq z\leq 5 \\
      0 & z\geq 5\\ 
\end{array} 
\right.
\end{equation}

where $H_0 = 70 \text{km s}^{-1}\text{Mpc}^{-1}$ is the local Hubble constant, and $\Omega_{\Lambda}=0.67$ is the universe's vacuum energy density.
Here we choose an observing period of $\Delta \tau_0 = 1$ year, and calculate the detection rate for the upper, central, and lower limits of the local binary NS coalescence rate density $\mathcal{R}=1540^{+3200}_{-1220} \text{ Gpc}^{-3}\text{yr}^{-1}$~\cite{Abbott2017}, giving the rates $N_A$ shown in the second column of Table~\ref{tab:variances}.

\item[(iii)] Compute the combined population standard deviation $\sigma_{N_A}$, taking into account sources at varying redshifts as was done in Eq. (3) of Ref.~\cite{Takahiro}:
\begin{equation}
\sigma_{N_A}^{-2}=\Delta \tau \int\limits^{z_h}_04 \pi \lbrack a_0 r_1(z)\rbrack^2 \mathcal{R}r(z)\frac{d\tau}{dz}\sigma^A_i(z)^{-2}dz.
\end{equation}
Here, we compute the populations' single-event uncertainties $\sigma_i^A(z)$ via the simple SNR scaling factor: 
\begin{equation}
\frac{\rho_i^A}{\rho_{\text{GW170817}}^A} = \frac{\sigma_{\text{GW170817}}^A}{\sigma_i^A},
\end{equation}
where $\rho_i^A$ is the SNR of simulated event $i$ computed from Eq.~\ref{eq:population}, and $\rho_{\text{GW170817}}^A$ and $\sigma_{\text{GW170817}}^A$ are the known SNRs and uncertainties of the GW170817 event from step (i).
This results in the combined statistical uncertainties for the lower, central, and upper limits of the local binary NS coalescence rate density $\mathcal{R}$, depicted by the cyan shaded region in Fig.~\ref{fig:stackedFisher}.
\end{enumerate}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%% APPENDIX B: MEASUREMENT ACCURACY AS A FUNCTION OF Q %%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Measurement accuracy as a function of binary mass ratio}\label{app:measurement}

In this appendix, we discuss the effects of the binary NS mass ratio $q$ on the measurement accuracy of $\tilde\Lambda$ for various interferometers. 
The top panel of Fig.~\ref{fig:massVariation} displays an approximation of the $\tilde\Lambda$ measurement accuracy as a function of increasing mass ratio $q$ (for a fixed chirp mass of $\mathcal{M}=1.188 \text{ M}_{\odot}$ corresponding to GW170817), using the Fisher analysis techniques described in Sec.~\ref{sec:observations}.
We observe two points of interest about these trends: (i) the second generation detectors (O2, aLIGO, A\texttt{+}, and Voyager) typically become less accurate as the mass ratio approaches unity; and (ii) the third generation detectors (CE and ET) observe the opposite behavior-- a fast growth in accuracy as the mass ratio increases.
For the remainder of this appendix, we attempt to explain the opposing behaviors of the second and third generation interferometers on the extraction of $\tilde\Lambda$.

\begin{figure}
\begin{center} 
\includegraphics[width=.3\textwidth]{massVariation1.eps}
\includegraphics[width=.3\textwidth]{massVariation2.eps}
\includegraphics[width=.3\textwidth]{massVariation3.eps}
\end{center}
\caption{
Measurement accuracy of $\tilde\Lambda$ as a function of increasing mass ratio $q$ for interferometer sensitivities O2, aLIGO, A\texttt{+}, Voyager, CE, and ET.
Evaluated at a fixed chirp mass of $\mathcal{M}=1.188\text{ M}_{\odot}$ corresponding to GW170817, this is repeated for three cases: (i) (top) with correlations between all template waveform parameters $\theta^a$ intact; (ii) (center) with the correlations between $\tilde\Lambda$ and $\delta\tilde\Lambda$ removed; and (iii) (bottom) with correlations between all parameters removed.
The removal of parameter correlations is approximated by setting certain Fisher matrix elements to 0, as described in Appendix.~\ref{app:measurement}.
Observe how the top panel shows strong disagreement between second and third generation detectors, while the bottom panel shows identical behavior.
This indicates that parameter correlations between $\tilde\Lambda$ and higher PN order parameter $\delta\tilde\Lambda$ are the culprit in such a disagreement.
}
\label{fig:massVariation}
\end{figure}

We begin by considering the correlations between individual template parameters $\theta^a$ and $\tilde\Lambda$, given in Eq.~\ref{eq:template}.
A simple approximation for the amplitude of these correlations can be performed by assuming that all of the other correlations are non-existing.
Individual correlations between parameters $\tilde\Lambda$ and $\theta^a$ is achieved by setting all of the off-diagonal Fisher matrix elements to $0$, except for the $(i,j)$ components such that both $i$ and $j$ correspond to the $\tilde\Lambda$ and $\theta^a$ parameters.
We follow this process for each parameter in turn, and display the results in Fig.~\ref{fig:correlations} for each interferometer.
This figure, showing only the relative differences between correlations of different parameters $\theta^a$ and $\tilde\Lambda$, reveals a few things.
First, the coalescence time $t_c$ is highly correlated with $\tilde\Lambda$ equally for each interferometer, an unsurprising fact due to the high 4pN order at which the parameter enters the waveform.
Secondly, correlations with 6pN tidal parameter $\delta\tilde\Lambda$ shows high correlations with $\tilde\Lambda$ for only the third generation detectors.
This can be explained by the high pN order at which $\delta\tilde\Lambda$ enters the waveform, an effect that only the most sensitive of detectors can probe.

\begin{figure}
\begin{center} 
\begin{overpic}[width=\columnwidth]{correlations.eps}
\put(121,13){\fontsize{7pt}{7pt}\selectfont $\mathcal{M}$}
\end{overpic}
\end{center}
\caption{
Measurement accuracy of $\tilde\Lambda$ upon the consideration of correlations between \emph{only} $\tilde\Lambda$ and other individual template parameters $\theta^a$ shown in Eq.~\ref{eq:template}.
This process, repeated for 6 interferometer design sensitivities O2, aLIGO, A\texttt{+}, Voyager, CE, and ET, is approximated by setting all off-diagonal Fisher matrix elements to 0 except for the $(i,j)$ components such that both $i$ and $j$ correspond to the $\tilde\Lambda$ and $\theta^a$ parameters.
Observe how correlations with 4PN parameter $t_c$ (coalescence time) remain to be highly correlated with $\tilde\Lambda$ among all 6 interferometers.
However, correlations between 6PN tidal parameter $\delta\tilde\Lambda$ and $\tilde\Lambda$ only begin to make a strong impact for third generation detectors.
This is due to the small effect that $\delta\tilde\Lambda$ has on the waveform, only to be distinguished by the most sensitive detectors.
}
\label{fig:correlations}
\end{figure}

Following this conclusion on the difference of correlations between tidal parameters $\tilde\Lambda$ and $\delta\tilde\Lambda$ for second and third generation detectors, we attempt to describe the opposing behaviors of measurement accuracy between the two.
Shown in the central panel of Fig.~\ref{fig:massVariation}, we approximate the measurement accuracy on $\tilde\Lambda$ with the correlations between $\delta\tilde\Lambda$ and $\tilde\Lambda$ removed, through a process similar to that described above. 
Notice how the trends between detector generations become more similar -- implying that the correlations between tidal parameters may be the cause of the opposing behaviors.
This is confirmed by the bottom panel of Fig.~\ref{fig:massVariation} where all parameter correlations have been removed from the Fisher matrix.
We see that this results in identical behaviors between all 6 representative interferometers, leading us to conclude that the discrepancy between the interferometers originates from the degeneracies between different parameters in the gravitational waveform, namely the 6pN tidal parameter $\delta\tilde\Lambda$ which is only detectable by the highly sensitive third generation detectors CE and ET.

%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%
%%%% BIBLIOGRAPHY %%%%
%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%

%\clearpage
\bibliography{Zack}
\end{document}