### 2019-04-09 paper/Response.txt
writing response to referee, as well as editing the paper

Dear Editor,

We would like to thank the referee for carefully reading our manuscript 
and providing valuable comments which have helped us improve the quality 
of our paper. For ease, we have highlighted all of our newest updates in 
boldface black. Additionally, we have outlined all of the comments
below, with "R:" and "A:" corresponding to the referee's comments and
the authors' replies, respectively.

We hope you reconsider publication of our manuscript in your journal.

Sincerely,

Zack Carson, Katerina Chatziioannou, Carl-Johan Haster, Kent Yagi, and 
Nicolás Yunes

---------------------------------------
Comments from the Referee
---------------------------------------

R: 1. I find the description of the “constrained” prior and its relation
to the “unconstrained” one slightly confusing. Sec. I A suggests that
it is constructed by sampling a large number of spectral EOSs and then
discarding those lying outside the 90% credible region of the EOS
posterior from GW170817. On the other hand, Sec. II A says that it is
built up from EOSs sampled directly from the EOS posterior (itself
derived from the unconstrained prior?), with no mention of the 90%
credible region. Naively, I would think that one would simply take the
unconstrained prior and reweight it by the GW170817 EOS likelihood to
get the constrained prior. Please explain more clearly which procedure
was adopted, and emphasize how the constrained and unconstrained
priors are related.

A: We apologize for the loss in translation between the descriptions of 
the constrained EoSs in the executive summary and Sec IIa. Ultimately, 
the "constrained" EoSs were computed identically to how the 90% 
posterior EoS were found in the paper by Abbott et al. So essentially we
are sampling the actual 90% posterior given by LIGO (the exact 
posterior EoSs from the paper are not publicly available, so the 
computation had to be re-done). Then, the unconstrained EoSs were 
generated completely randomly, with no relation to the constrained EoS
or the posterior PDF. We have modified our descriptions of this in both 
Sec. Ia and IIa.

R: 2. To quantify the error in the tidal deformability introduced by the
approximately universal nature of the binary Love relations, Sec. IV
adopts a representation for the error about the best-fit relation. The
error model is fitted to data from the constrained prior. Does this
prior include any hybrid EOSs, the models that deviate the most from
the binary Love relations? If not, do the projections in Sec. IV rely
on an implicit assumption that the EOS is purely hadronic?

A: Yes, you are correct that our results for the systematic errors from
the binary Love relations rely on the assumption of hadronic stars. Our 
set of constrained EoSs do not contain any hybrid star EoSs, following
from the calculations of Abbott et al on the EoS posterior probability 
distribution, which also assumed purely hadronic EoSs. If included these
results may change slightly, and could be a possible avenue of future
work. We have added a couple of sentences to the end of Sec. IVa 
noting this.

R:  3. Several papers have claimed that EOSs with sharp phase transitions
can violate I-Love-Q universality, e.g. Han & Steiner,
arXiv:1810.10967; Bandyopadhyay et al., Eur. Phys. J. A (2018) 54: 26;
Lau et al., Phys. Rev. D 95, 101302 (2017). Here, in Sec. III, hybrid
hadron-quark EOSs are found to obey the standard I-Love-Q relations.
Is this finding compatible with the existing claims in the literature?
The paper would benefit from a brief discussion of this point.

A: Upon reading over the papers by Han & Steiner, Bandyopadhyay et al, 
and Lau et al, we find that our results are indeed consistent with what
they present. 
The paper by Han & Steiner claims that "for typical neutron
stars > 1.0 M_sun, discrepancies are negligible", which can also be seen
in the bottom left panel of Fig. 11. In fact, the fits are only valid in
the M>1M_sun regime as can be seen in the top panel of Fig. 11, a relic 
fixed in our paper by using improved fits which limit to the Newtonian
case.
In the paper by Lau et al,  they claim that the "I-Love 
relation for hybrid stars still agrees with the universal relation for 
fluid stars to about 1% level for hybrid stars with a quark-matter core
surrounded by a fluid [hadronic] envelope", similar to what we investigate. 
They find that solid quark-matter stars (not studied by us) violate the 
relations proportionally to the value of the gap parameter. 
Finally, in the paper by Bandyopadhyay et al, there is no claim to *how* 
strongly the "HQ" EoSs violate the I-Love relation. However, by looking 
at Fig. we see that the departure from the "np" EoSs is not noticeably 
large, and seems to be somewhat consistent with what we see with our 
hybrid star EoSs. 
We also see in Paschalidis2018 (10.1103/physrevd.97.084038) that it was 
shown that the I-Love-Q universality is preserved for the hybrid stars
considered, which are identical to the ones used here.
As so, we have added a couple of sentences at the end of Sec.
IIIA2 discussing the consistency, as well as the additional references.

R:  4. In Fig. 5, the departure of the hybrid EOSs from the best-fit
binary Love relation appears to become more pronounced as q tends to
1. Is there an intuitive explanation for this behavior?

A: This behavior is caused for the same reason as the (fractional) error 
increases for hadronic stars with increasing mass ratio. This is because 
as the mass ratio decreases, the universal relation becomes more exact 
as Lambda_s -> Lambda_a (exactly equal in the q=0 limit). Therefore, as 
the mass ratio increases to 1, Lambda_a and Lambda_s become maximally 
different, and the fractional error becomes maximal. We have added a 
comment on this at the end of the caption on Fig. 5, as well as added 
the absolute differences for hybrid stars in the bottom three panels.

R:  * Sec. I: A more up-to-date reference about systematic errors in
neutron star mass-radius measurements is Miller & Lamb, F. Eur. Phys.
J. A (2016) 52: 63.

A: We thank the referee for the updated citation on M-R systematics. In 
response, we have added the Miller & Lamb 2016 citation in the first 
paragraph of Sec. I.

R: * Sec. I: The last sentence of the second paragraph seems to attribute
the tidal signature in the waveform solely to the modified orbital
dynamics, but there is an equally important dissipative contribution
due to the extra multipole radiation sourced by the tidal bulges.

A: We have modified the sentence in question to include this additional
GW modification effect: "These deformations source additional multipole radiation 
as well as affect the orbital trajectory of the binary, thus altering the
 GWs emitted, encoding within the latter the NS EoS."

R: * Sec. I A: Surely the variability in the R-Love relation is on the
order of hundreds of meters, not hundreds of kilometers.

A: We thank the referee for finding this typo! It is indeed errors of
~880 m and ~360 m, not km. This has been adjusted accordingly.

R: * Sec. II A: Why is causality only required to within 10% in the
prior? It does not appear to be explained in the text. (Also, Gamma is
incorrectly called the polytropic index instead of the adabiatic index
in this discussion.)

A: We chose causality to 10% to be consistent with the analysis of 
Abbott et al. 2018 (10.1103/PhysRevLett.121.161101), from which they 
choose to adhere to causality up to 10% upon deriving the EoS 
posterior probability distribution. We have added a short note to Sec. 
IIa. commenting on this. We have also fixed the incorrect naming of Gamma.

R: * Sec. IV: The sentence reporting the GW170817 Lambda_tilde
measurement does not cite the reference from which it was taken.
(Also, there appears to be an extraneous comment embedded in the same
sentence.)

A: We thank the referee for finding the missing citation and extraneous 
comment that somehow made it into the final draft. The citation has been
added, and comment removed.

---------------------------------------
Other Modifications
---------------------------------------

- We have included additional citations regarding the sentence 
"...but unfortunately, these [modified theories of gravity] are 
typically degenerate with the nuclear EoS. These are Pani & Berti 2014, 
Arxiv:1405.4547, Minamitsuji & Silva 2016, Arxviv:1604.07742, and 
Maselli & Silva 2016, Arxiv:1603.04876